### 2016-05-09 Article/main.tex
Adding smartphone journal paper

\documentclass[conference]{IEEEtran}
% setup page to suit conference specification using fancyhdr
\usepackage{fancyhdr}
\setlength{\paperwidth}{215.9mm}
\setlength{\hoffset}{-9.7mm}
\setlength{\oddsidemargin}{0mm}
\setlength{\textwidth}{184.3mm}
\setlength{\columnsep}{6.3mm}
\setlength{\marginparsep}{0mm}
\setlength{\marginparwidth}{0mm}

\setlength{\paperheight}{279.4mm}
\setlength{\voffset}{-7.4mm}
\setlength{\topmargin}{0mm}
\setlength{\headheight}{0mm}
\setlength{\headsep}{0mm}
\setlength{\topskip}{0mm}
\setlength{\textheight}{235.2mm}
\setlength{\footskip}{12.4mm}

\setlength{\parindent}{1pc}

% If IEEEtran.cls has not been ins

\usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{subcaption}
\usepackage{multicol}

\hyphenation{op-tical net-works semi-conduc-tor}

 \setlength{\columnsep}{1.0cm}

\begin{document}

 \title{\fontsize{24pt}{10pt}\selectfont\textbf{OpenMotion}: An open source library for attitude estimation.}

\author{\IEEEauthorblockN{Nizar Ouarti}
\IEEEauthorblockA{IPAL (Sorbonne UPMC, CNRS, \\Astar, NUS, UJF, IMT)\\
Singapore\\
Email: nizar.ouarti@ipal.cnrs.fr}
\and
\IEEEauthorblockN{Thomas Braud}
\IEEEauthorblockA{IPAL (Sorbonne UPMC, CNRS,\\ Astar, NUS, UJF, IMT)\\
Singapore\\
Email: braud.thomas@ipal.cnrs.fr }
\and
\IEEEauthorblockN{Vivien Billaud}
\IEEEauthorblockA{IPAL (Sorbonne UPMC, CNRS,\\ Astar, NUS, UJF, IMT)\\
Singapore }}

\maketitle

\thispagestyle{fancy}
\fancyhead{}
\lhead{}
\lfoot{}
\cfoot{}
\rfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\def\mathbi#1{\textbf{\em #1}}
\makeatletter
\newcommand*{\rom}[1]{\expandafter\@slowromancap\romannumeral #1@}
\makeatother

\begin{abstract}
The abstract goes here
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}

\section{Global Presentation of the Library}

\begin{itemize}
\item The North East Down (NED) frame $\{a\}$ system has its origin fixed at the (moving) object center of gravity. The $z$-axis points upward perpendicularly to the tangent plane of the ellipsoid, and the $x$-axis points towards true north (and not the magnetic north). The $y$-axis point towards east.

\vspace{0.1cm}

\item The object-fixed reference frame $\{c\}$ is a moving and rotating coordinate frame that is fixed to the device. The $x$-axis points in the forward direction, the $y$-axis to the right side and the $z$-axis upward. 
\end{itemize}

We admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure~\ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal} .

\begin{figure}[!h]
\centering
\includegraphics[scale=0.55]{images/Schema_situation.png}
\caption{Definition of the scene with 2 frame:  a fixed frame NED (North East Down) noted $\{a\}$ and a moving frame object noted $\{c\}$}
\label{Schema_situation}
\end{figure}

\section{Calibration and initialization}

\section{State of the art}

\subsection{Kinematic and quaternion representation}

A quaternion is a four-dimensional vector, defined as $ \mathbi{q} = \begin{pmatrix} \textbf{q} \\ q_w \end{pmatrix} $ with $ \textbf{q} =  ( q_x, q_y, q_z)^T $. For attitude representation, the quaternion must satisfies a single constraint given by  $|| \mathbi{q} || = 1$.The quaternion kinematics equation in discrete-time is given by :

\begin{equation}
\mathbi{q}_{k+1} = \Phi_{k+1|k}(\boldsymbol\omega)\mathbi{q}_{k}
\label{quat_kine}
\end{equation}

where $\boldsymbol\omega$ is the angular velocity and $\Phi_{k+4|k}$ is the transition matrix from time $k-1$ to $k$ such as:

\begin{equation}
 \Phi_{k+1|k}(\boldsymbol\omega) = \begin{pmatrix}  \cos(0.5||\boldsymbol\omega||\Delta T)I_3 - [\boldsymbol \psi\times] & \boldsymbol\psi \\ -\boldsymbol\psi^T &  \cos(0.5||\boldsymbol\omega||\Delta T)  \end{pmatrix} 
\end{equation}

 with,

\begin{equation}
 \boldsymbol \psi = \sin(0.5||\boldsymbol\omega||\Delta T)\frac{\boldsymbol\omega}{||\boldsymbol\omega||}
\end{equation}

And $[\textbf{v} \times] \in \mathbb{R}^{3\times 3}$ is the skew-symetric matrix of the vector $\textbf{v}$ such as.

\begin{equation}
[\textbf{v} \times] = \begin{pmatrix} 0 & v_z & -v_y \\ -v_z & 0 & v_x \\ v_y & -v_x & 0 \end{pmatrix} 
\label{skewsymmat}
\end{equation}
Then, the attitude matrix $R \in SO(3) $ related to the quaternion $\mathbi{q}$ is given by :

\begin{equation}
R(\mathbi{q}) = (q_w^2 - |\textbf{q}|^2)I + 2 \textbf{q}\textbf{q}^T - 2q_w[\textbf{q} \times]
\label{quat_to_rot}
\end{equation}

Where $SO(3)$ is the special orthogonal group defined by:

\begin{center}
$SO(3) = \{ R \in \mathbb{R}^{ 3 \times 3} | RR^T = R^TR = I, det(R) = 1 \}$
\end{center}

\subsection{Most Common Fusion Methods}

\subsubsection{Kalman Filtering Methods}

Introduced in 1960 by Kalman \cite{kalman_new_1960}, the Kalman filter is an algorithm that uses a series of measurements observed over time, containing noise, in order to estimate unknown variables (state vector). %The goal is to obtain more precise results than those based on a single measurement alone. 
Based on a model it is possible to obtain a more robust estimation of the state vector. Consider the following linear system, described by the difference equation and the observation model with additive noise:

\begin{equation}
\left\{ \begin{array}{l}
\textbf{x}_{k} = F_{k}\textbf{x}_{k-1}+B_{k}\textbf{u}_{k}+G_{k}\textbf{v}_{k}\\
\textbf{z}_{k} = H_{k}\textbf{x}_{k}+D_{k}\textbf{w}_{k}
\end{array} \right.
\label{linear_system}
\end{equation}

with,

\begin{center}
\begin{tabular}{rl}
$\textbf{x}_{k} $ &State vector at time $k$\\
$\textbf{z}_{k} $ &Measurement vector at time $k$\\
$\textbf{u}_{k} $ & Input vector (e.i.: from sensor) at time $k$ \\
$\textbf{v}_{k} $ & State noise vector disturbing the system\\
$\textbf{w}_{k} $ & Observation noise disturbing the measurement\\
\end{tabular}
\end{center}

\vspace{0.2cm}

The figure \ref{State_equation_discrete_case} represents the dynamical system described in \ref{linear_system} . A brief presentation of the algorithm is provided here,  see \cite{terejanu2013discrete} for more details. The algorithm has 3 important steps :\\

\begin{itemize}
\item The first step is the \textbf{prediction} of state $\textbf{x}^-_{k}$ at time $k$ and covariance $P^-_{k}$ is given by :

\begin{equation}
\left\{ \begin{array}{cl}
\textbf{x}^-_{k} = & F_{k}\textbf{x}^+_{k-1} + B_{k}\textbf{u}_{k} \\
P^-_{k} = & F_{k}P^+_{k-1}F^T_{k}+G_{k}Q_{k}G^T_{k}
\end{array}
\right.
\end{equation}

\item The \textbf{innovation} is the difference between predicted measurement and real measurement. It measure up the deviation used by the Kalman gain.
\begin{equation}
\left\{ \begin{array}{cl}
\boldsymbol\nu_{k} = & \textbf{z}_{k}-H_{k}\textbf{x}^-_{k} \\
S_{k} = & R_{k} +H_{k}P^-_{k}H^T_{k}
\end{array}
\right.
\end{equation}

\item And finally, An \textbf{update} is processed to correct the estimate state $\textbf{x}^+_{k} $ and his covariance $P^+_{k}$ with the innovation.

\begin{equation}
\left\{ \begin{array}{cl}
\textbf{x}^+_{k} = & \textbf{x}^-_{k} + K_{k}\boldsymbol\nu_{k} \\
P^+_{k} = &P^-_{k} - K_{k}S_{k}K^T_{k}
\end{array}
\right.
\end{equation}

\end{itemize}

Where $K_k$ is the Kalman gain  representing relative importance of innovation  $\nu_{k}$.

\begin{equation}
K_{k}=P^-_{k}H^T_{k}S^{-1}
\end{equation}

%And so on for future state $k+1,...$.

\subsubsection{Nonlinear Observer}

The theory of the state observer was first introduced by Kalman and Bucy for a linear system in a stochastic environment. Then Luenberger \cite{david1971introduction} made a general theory of observers for deterministic linear systems, introducing the notions of observer and reduced minimum observer \cite{primbs1996survey}. Mahony et al. \cite{mahony_nonlinear_2008} have proposed an nonlinear observer called Constant Gain Observer (CGO), termed the explicit complementary filter, that provides attitude estimates as well as gyro bias estimates.

\begin{itemize}
\item Prediction
\end{itemize}

First, we propagate the estimate attitude and the gyro bias with equation \ref{prediction_cgo}.

\begin{equation}
\left\{\begin{array}{l}
\mathbi{q}_{k}^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro} - \boldsymbol \beta_{k-1}^+)\mathbi{q}_{k-1}^+ \\
\boldsymbol \beta_k^- = \boldsymbol \beta_{k-1}^+
 \end{array}
\right.
\label{prediction_cgo}
\end{equation}

\begin{itemize}
\item Correction
\end{itemize}

The idea is to compute a  constraint  given by a comparison between  an estimate value of a sensor and his real data at a time $k$.

\begin{equation}
\left\{\begin{array}{l}
\textbf{v}_{acc}^- = R_(\mathbi{q}_{k}^-)\textbf{a}^a + \boldsymbol  \beta_{acc} \\
\textbf{v}_{mag}^- = R_(\mathbi{q}_{k}^-)\textbf{m}^a + \boldsymbol  \beta_{mag}
 \end{array}
\right.
\end{equation}

All these vectors are normalized.

The '' gain '' is given by,

\begin{equation}
\boldsymbol \omega_{mes} = -vex(\sum_{i=1}^n \frac{k_i}{2}(\textbf{v}_i(\textbf{v}_i^-)^T -  \textbf{v}_i^-(\textbf{v}_i)^T )) 
\label{cgo_gain}
\end{equation}

where the $vex$ operator define by the inverse of the cross product.
\begin{equation}
\left\{\begin{array}{l}
\textbf{a}(\textbf{b})^T -  \textbf{b}(\textbf{a})^T  = [ (\textbf{a} \times \textbf{b}) \times ]\\
vex([\textbf{a}\times])  = \textbf{a}
 \end{array}
\right.
\end{equation}

Then, we the gain is used to correct the predicted attitude $\mathbi{q}_{k}^-$ by the following equation:

\begin{equation}
\left\{\begin{array}{l}
\mathbi{q}_{k}^+ = \mathbi{q}_{k}^- + \frac{\Delta T}{2} \Xi(\mathbi{q}_{k}^-) \boldsymbol\omega_{mes}k_p\\
\boldsymbol \beta_k^+ = \boldsymbol \beta_{k}^-  - \boldsymbol\omega_{mes}k_i
 \end{array}
\right.
\end{equation}

One problem of Kalman Filters is to find the initial values in the covariance (matrix $Q$) because it is difficult to relate it to real physical quantities. The covariance matrix is not involve in nonlinear observer theory. This is one of the reasons there has been an increasing interest for nonlinear observer design the last decades. Another reason is their low computational need.

\subsubsection{Methods Based on Solutions of the Wahba's Problem}

The attitude determination problem for vector observation was first formulated as a least squares estimation problem by Wahba \cite{wahba_least_1965} in 1965 :\\

Given the two sets of $n$ vectors $\{ \textbf{r}_{1},...,\textbf{r}_{n} \}$ and $\{ \textbf{b}_{1},...,\textbf{b}_{n} \}$, $n \geqslant 2 $, where each pair $(\textbf{r}_{i},\textbf{b}_{i})$ corresponds to a generalised vector, $\textbf{x}_{i}$, find the proper orthogonal matrix, $A$, which bring the first into the best least squares coincidence with the second. That is, find $A$ which minimises the function $J$ defined as:

\begin{equation}
J(A) = \frac{1}{2} \sum_{i=1}^n a_i||b_i-Ar_i||^2
\label{Wahba_loss_function}
\end{equation}

subject to the constraint $A^TA = I$ and $det(A)=1$. The equation (\ref{Wahba_loss_function}) can be written with a quaternion representation which gives:

\begin{equation}
J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i||b_i-R(\mathbi{q})r_i||^2
\label{QUEST}
\end{equation}

The matrix $R(\mathbi{q})$, is the rotation matrix that corresponds to the quaternion $\mathbi{q}$ (equation \ref{quat_to_rot}). The QUEST algorithm are conducted in the following manner. Let

\begin{equation}
B = \sum_{i=1}^N a_i\textbf{b}_i \textbf{r}_i^T
\end{equation}

Then, we can rewrite the equation (\ref{QUEST}) as follow

\begin{equation}
J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i  - tr(R(\mathbi{q})B^T)
\label{QUEST_bis}
\end{equation}

Due to the fact that the representation of the attitude matrix is a homogenous quadratic function of $\mathbi{q}$, we can say that

\begin{equation}
tr(R(\mathbi{q})B^T) = \mathbi{q}^TK\mathbi{q}
\label{QUEST_ter}
\end{equation}

where $K$ is the symmetric traceless matrix such as

\begin{equation}
K = \begin{pmatrix} S-tr(B)I_3 & \textbf{z} \\ \textbf{z}^T & tr(B)
\end{pmatrix}
\end{equation}

with

\begin{equation}
\left\{\begin{array}{l}
S = B + B^T\\
\textbf{z} = \sum_ia_i\textbf{b}_i\times\textbf{r}_i
 \end{array}
\right.
\end{equation}

The minimisation problem from Equation (\ref{QUEST_bis}) is now a maximisation problem. thus, the solution $\mathbi{q}_{opt}$ is the eingen vector correspond to the largest eigen value of $K$ \cite{markley1999estimate}. The main problem of QUEST is that it does not take into account of all the previous observations performed since the beginning. A recursive QUEST algorithm (REQUEST) proposed by Bar-Itzhack manages this problem. According to the REQUEST algorithm, the propagation of $K_k$ is given by

\begin{equation}
K_k^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro})K_{k-1}^+\Phi_{k|k-1}^T(\boldsymbol\omega_{gyro})
\end{equation}

We compute $\delta K_k $ as

\begin{equation}
\delta K_k = \frac{1}{\sum_ia_i}\begin{pmatrix} \delta S_k-\delta\sigma_kI_3 & \delta\textbf{z}_k \\ \delta\textbf{z}_k^T & \delta\sigma_k
\end{pmatrix}
\end{equation}

with,

\begin{equation}
\begin{array}{cc}
\delta B_k = \sum_i a_i \textbf{b}_i\textbf{r}_i^T & \delta S_k = B_k + B_k^T\\
\delta \textbf{z}_k = \sum_ia_i\textbf{b}_i\times\textbf{r}_i & \delta \sigma_k = tr(B_k)
\end{array}
\end{equation}

Where $\textbf{b}_i$ is the observation vector at time $k$. Then we update the K-matrix.

\begin{equation}
K_k^+ = \frac{\rho_k m_{k-1}}{m_k}K_k^- + \frac{1}{m_k}\delta K_k
\end{equation}

A drawback is that the value of $\rho_k$ comes from heuristics. Choukroun revisit REQUEST provides an improvement with his Optimal REQUEST \cite{choukroun_novel_2003} where an optimal $\rho_k^*$ is computed which gives better results. Here we choose to implement this version of REQUEST. Contrary to other methods, the QUEST algorithms do not estimate the gyroscope bias. Which can make it less robust in some cases.
There are also other methods based on Wahba's problem like the Davenport's q-method \cite{weighted1971nasa} proposed in 1971 considered more robust than QUEST. Or the method ESOQ proposed by Mortari \cite{mortari1997esoq}  in 1997. But in this study we limit our investigation to REQUEST.\\

\subsubsection{Particle Filtering Method}

Particle filtering method is based on Monte-Carlo method\cite{metropolis1949monte}. The main idea is to use a large number of samples called particles to estimate a nonlinear function \cite{chen_bayesian_2003} and \cite{terejanu2009tutorial}. The set of $N$ particle with associated weight at time $k$ are denoted by $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$. Initially, the particle are drawn from a proposal distribution. The PF is divided into three step, prediction, update, resampling, that constitute a filter cycle:

\begin{itemize}
\item Prediction
\end{itemize}

The first step is to propagate all particles $\{\textbf{x}^{(i)}_{k-1},w^{(i)}_{k-1}\}$ to $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$ with the nonlinear function $f$ while the weight remain unchanged :

\begin{equation}
\textbf{x}^{(i)}_{k} = f(\textbf{x}^{(i)}_{k-1},\textbf{u}_{k-1}) + \textbf{w}_{k-1}^{(i)}\\
\end{equation}

where $N$ sample $\textbf{w}_{k-1}^{(i)}$ of the process noise (supposed gaussian in our case) are drawn.

\begin{itemize}
\item Update/Correction
\end{itemize}

At the update step, the weights associated with each particle is updated based on the likelihood function $p( \textbf{z}_k| \textbf{x}_k^{(i)})$.
\begin{equation}
\left\{ 
\begin{array}{l}
w_k^{(i)} = w_{k-1}^{(i)} p( \textbf{z}_k| \textbf{x}_k^{(i)})\\
\tilde{w}_k^{(i)} = \frac{w_{k}^{(i)}}{\sum_iw_{k}^{(i)}}
\end{array}
\right.
\end{equation}

The weight are normalised such as we have $\sum_i\tilde{w}_k^{(i)}=1$. To finish, the mean is computed with the following equation:

\begin{equation}
 \textbf{x}_k^+ = \sum_{i=1}^N\tilde{w}_k^{(i)} \textbf{x}_k^{(i)}
 \end{equation}

\begin{itemize}
\item Resampling/Regularisation
\end{itemize}

A common problem with PF is the degeneracy phenomenon. After some iterations, more and more particles will have a weak weight, which make them negligible and useless.  To overcome this problem, a regularisation step called resampling is processed. It is not needed to perform PF but it allows to maintain a good performance relative to degeneracy problem. The idea of resampling is to eliminate particles having feeble weight and focus on strong weight particles. In order to measure degeneracy, the number of effective samples $N_{eff}$ is computed (equation \ref{N_eff}). If  $N_{eff}$ is lower than a threshold, the resampling step is processed.

\begin{equation}
 N_{eff} = \sum_{i=1}^N\frac{1}{(\tilde{w}_k^{(i)})^2}
 \label{N_eff}
 \end{equation}

The choice of the threshold is important. Indeed, a high threshold gives $N_{eff} \approx N$, but that does not mean necessarily we have a lot of efficient particles. On the contrary, it means that most of the particle are identical making large computations unnecessary. Moreover, with a low threshold, many particles are neglected making computations less efficient. In general, the threshold is fixed at  $\frac{N}{2}$ or $\frac{N}{3}$.\\

There are several types of resampling, a comparison is made in \cite{douc2005comparison}. We took the resampling algorithm described in \cite{arulampalam2002tutorial} because it is the most commun. Cheng, Yang and Crassidis have propose an attitute estimation with a particle filter method \cite{cheng_particle_2010}. We have also considered the revisited algorithm of Cheng \cite{chang_particle_2014} based on bootstrap filter\cite{gordon1993novel}. We have chosen to implement this one because they use the same approach as USQUE \cite{crassidis_unscented_2003} and MEKF\cite{markley2003attitude}. We decided to explore the performance of this three algorithms with the same model. \\

The particle filter provides an estimate of the nonlinear system without assuming that the noise is Gaussian. So it is more robust to any situation, which makes it more performant in the context of a real applications. But an important computation time is required and that is a limitation for a real-time application.

\section{Performance Evaluation}

We have access through our designed ground truth to the true attitude $\mathbi{q}^{real}_k$ at time $k$ of the reference frame $\{c\}$ related to $\{a\}$.
By computation, each method returns an estimate attitude $\mathbi{q}^{est}_k$ at time $k$ (see figure \ref{test_method}).

\begin{equation}
 \delta \mathbi{q}_k = \mathbi{q}^{real}_k \otimes (\mathbi{q}^{est}_k)^{-1} =  \begin{pmatrix}  \delta \textbf{q}_{k} \\ \delta q_{w_k}  \end{pmatrix}
\end{equation}

The error-MRP vector related to $ \delta \mathbi{q}_k$ is given by

\begin{equation}
 \delta \textbf{p}_k  = 4[\frac{\delta \textbf{q}_{k}  }{(1+\delta q_{w_k})}]
\end{equation}

Thus, we can calculate the attitude error given by the root mean square (RMS) of $\delta \textbf{p}_k$:

\begin{equation}
\epsilon_k  = RMS (\delta \textbf{p}_k) = \sqrt{\frac{1}{3}( \delta p_{k_x}^2 +\delta p_{k_y}^2+\delta p_{k_z}^2)  }
\label{error_definition}
\end{equation}

The mean of the error is given by :

\begin{equation}
\bar{\epsilon}  = \frac{1}{n}\sum_{i=1}^n \epsilon_i \\
\label{mean_var_error_def}
\end{equation}

With $n$ the number of data at our disposal. We use this information (mean error) for our performance study.

\begin{figure}[!h]
\centering
\includegraphics[scale=0.40]{images/test_method.png}
\caption{Algorithm designed to compute attitude estimation methods' performances}
\label{test_method}
\end{figure}

To differentiate the methods, we assigned  some ''scores'' in order to highlight the behaviour of these methods in different configurations. We defined these ''scores'' as:

\begin{align}
s_1 = \bar{\epsilon}\\
s_2 = \frac{1}{\bar{\epsilon} \times \bar{\tau}}
\label{score}
\end{align}

With $\bar{\epsilon}$ average attitude error (voir \ref{error_definition}) et $\bar{\tau}$ the average computation time. \\

\begin{figure}[!h]

\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.85\linewidth]{images/histo_s1_add.png}
  \caption{Mean Attitude Error (score $s_1$)}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.85\linewidth]{images/histo_s2_add.png}
  \caption{Ratio Error times Computation Time (score $s_2$)}

\end{subfigure}
\caption{Scores of methods where noise is AWGN with 3 strength}

\end{figure}

\section{Conclusion}

notre library est trop bien...

\bibliographystyle{ieeetr}

\end{document}

### 2016-05-09 Article/main.tex
adding smartphone journal template

@@
-\documentclass[conference]{IEEEtran}
-% setup page to suit conference specification using fancyhdr
-\usepackage{fancyhdr}
-\setlength{\paperwidth}{215.9mm}
-\setlength{\hoffset}{-9.7mm}
-\setlength{\oddsidemargin}{0mm}
-\setlength{\textwidth}{184.3mm}
-\setlength{\columnsep}{6.3mm}
-\setlength{\marginparsep}{0mm}
-\setlength{\marginparwidth}{0mm}
+\documentclass[twocolumn]{bmcart}
 
-\setlength{\paperheight}{279.4mm}
-\setlength{\voffset}{-7.4mm}
-\setlength{\topmargin}{0mm}
-\setlength{\headheight}{0mm}
-\setlength{\headsep}{0mm}
-\setlength{\topskip}{0mm}
-\setlength{\textheight}{235.2mm}
-\setlength{\footskip}{12.4mm}
-
-\setlength{\parindent}{1pc}
-
-% If IEEEtran.cls has not been ins
+\usepackage[utf8]{inputenc} %unicode support
 
@@
 \usepackage{subcaption}
-\usepackage{multicol}
 
-\hyphenation{op-tical net-works semi-conduc-tor}
-
- \setlength{\columnsep}{1.0cm}
-
-\begin{document}
-
- \title{\fontsize{24pt}{10pt}\selectfont\textbf{OpenMotion}: An open source library for attitude estimation.}
-
-\author{\IEEEauthorblockN{Nizar Ouarti}
-\IEEEauthorblockA{IPAL (Sorbonne UPMC, CNRS, \\Astar, NUS, UJF, IMT)\\
-Singapore\\
-Email: nizar.ouarti@ipal.cnrs.fr}
-\and
-\IEEEauthorblockN{Thomas Braud}
-\IEEEauthorblockA{IPAL (Sorbonne UPMC, CNRS,\\ Astar, NUS, UJF, IMT)\\
-Singapore\\
-Email: braud.thomas@ipal.cnrs.fr }
-\and
-\IEEEauthorblockN{Vivien Billaud}
-\IEEEauthorblockA{IPAL (Sorbonne UPMC, CNRS,\\ Astar, NUS, UJF, IMT)\\
-Singapore }}
-
-\maketitle
-
-\thispagestyle{fancy}
-\fancyhead{}
-\lhead{}
-\lfoot{}
-\cfoot{}
-\rfoot{}
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrulewidth}{0pt}
+\startlocaldefs
 
@@
 
-\begin{abstract}
+\endlocaldefs
+
+%%% Begin ...
+\begin{document}
+
+%%% Start of article front matter
+\begin{frontmatter}
+
+\begin{fmbox}
+\dochead{Research}
+
+\author[
+   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
+   corref={aff1},     
+   noteref={n1},                        % id's of article notes, if any
+   email={nizar.ouarti@ipal.cnrs.fr}   % email address
+]{\inits{NO}\fnm{Nizar} \snm{Ouarti}}
+\author[
+   addressref={aff1},   corref={aff1},  
+   email={braud.thomas@ipal.cnrs.fr}
+]{\inits{TB}\fnm{Thomas} \snm{Braud}}
+\author[
+   addressref={aff1},   corref={aff1},  
+   email={vivien.billaud@ipal.cnrs.fr}
+]{\inits{VB}\fnm{Vivien} \snm{Billaud}}
+
+\address[id=aff1]{%                           % unique id
+  \orgname{IPAL (Sorbonne UPMC, CNRS, Astar, NUS, UJF, IMT)}, % university, etc
+  \city{Singapore}
+  }
+
+\end{fmbox}% comment this for two column layout
+
+ \title{OpenMotion: An open source library for attitude estimation.}
+
+\maketitle
+
+\begin{abstractbox}
+
+\begin{abstract} % abstract
+
 The abstract goes here
@@
 
-\IEEEpeerreviewmaketitle
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% The keywords begin here                  %%
+%%                                          %%
+%% Put each keyword in separate \kwd{}.     %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{keyword}
+\kwd{IMU}
+\kwd{Sensor Fusion}
+\kwd{Sensor Calibration}
+\end{keyword}
+
+\end{abstractbox}
+
+\end{frontmatter}
 
@@
 \section{Global Presentation of the Library}
+
+\texttt{OpenMotion} est une librairie open source qui permet a l'utilisateur d'estimer l'orientation du telephone a partir des donnee de l'IMU.
+La librairie ne supporte pas encore les translations, nous nous sommes concentre sur l'attitude.
 
@@
 
-\item The object-fixed reference frame $\{c\}$ is a moving and rotating coordinate frame that is fixed to the device.
-The $x$-axis points in the forward direction, the $y$-axis to the right side and the $z$-axis upward.
+\item The object-fixed reference frame $\{c\}$ is a moving and rotating coordinate frame that is fixed to the smartphone device. 
 \end{itemize}
 
-We admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure~\ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal} .
+The library admit admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure~\ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal} .
 
@@
 
+La librairie donne une estimation le quaternion $\mathbi{q}_a^c$
+
 \section{Calibration and initialization}
@@
 
-A quaternion is a four-dimensional vector, defined as $ \mathbi{q} = \begin{pmatrix} \textbf{q} \\ q_w \end{pmatrix} $ with $ \textbf{q} =  ( q_x, q_y, q_z)^T $.
+A quaternion is a four-dimensional vector, defined as
+
+\begin{center}
+$ \mathbi{q} = \begin{pmatrix} \textbf{q} \\ q_w \end{pmatrix} $ with $ \textbf{q} =  ( q_x, q_y, q_z)^T $. 
+\end{center}
+
 For attitude representation, the quaternion must satisfies a single constraint given by  $|| \mathbi{q} || = 1$.The quaternion kinematics equation in discrete-time is given by :

### 2016-05-09 Article/calibration.tex
adding different file for article

--- 2016-05-09 Article/main.tex
@@
-\documentclass[twocolumn]{bmcart}
-
-\usepackage[utf8]{inputenc} %unicode support
-
-\usepackage{graphics} % for pdf, bitmapped graphics files
-\usepackage{epsfig} % for postscript graphics files
-%\usepackage{mathptmx} % assumes new font selection scheme installed
-%\usepackage{times} % assumes new font selection scheme installed
-\usepackage{amsmath} % assumes amsmath package installed
-\usepackage{amssymb}  % assumes amsmath package installed
-\usepackage{subcaption}
-
-\startlocaldefs
-
-\def\mathbi#1{\textbf{\em #1}}
-\makeatletter
-\newcommand*{\rom}[1]{\expandafter\@slowromancap\romannumeral #1@}
-\makeatother
-
-\endlocaldefs
-
-%%% Begin ...
-\begin{document}
-
-%%% Start of article front matter
-\begin{frontmatter}
-
-\begin{fmbox}
-\dochead{Research}
-
-\author[
-   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
-   corref={aff1},     
-   noteref={n1},                        % id's of article notes, if any
-   email={nizar.ouarti@ipal.cnrs.fr}   % email address
-]{\inits{NO}\fnm{Nizar} \snm{Ouarti}}
-\author[
-   addressref={aff1},   corref={aff1},  
-   email={braud.thomas@ipal.cnrs.fr}
-]{\inits{TB}\fnm{Thomas} \snm{Braud}}
-\author[
-   addressref={aff1},   corref={aff1},  
-   email={vivien.billaud@ipal.cnrs.fr}
-]{\inits{VB}\fnm{Vivien} \snm{Billaud}}
-
-\address[id=aff1]{%                           % unique id
-  \orgname{IPAL (Sorbonne UPMC, CNRS, Astar, NUS, UJF, IMT)}, % university, etc
-  \city{Singapore}
-  }
-
-\end{fmbox}% comment this for two column layout
-
- \title{OpenMotion: An open source library for attitude estimation.}
-
-\maketitle
-
-\begin{abstractbox}
-
-\begin{abstract} % abstract
-
-The abstract goes here
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The keywords begin here                  %%
-%%                                          %%
-%% Put each keyword in separate \kwd{}.     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{keyword}
-\kwd{IMU}
-\kwd{Sensor Fusion}
-\kwd{Sensor Calibration}
-\end{keyword}
-
-\end{abstractbox}
-
-\end{frontmatter}
-
-\section{Introduction}
-
-\section{Global Presentation of the Library}
-
-\texttt{OpenMotion} est une librairie open source qui permet a l'utilisateur d'estimer l'orientation du telephone a partir des donnee de l'IMU.
-La librairie ne supporte pas encore les translations, nous nous sommes concentre sur l'attitude.
-
-\begin{itemize}
-\item The North East Down (NED) frame $\{a\}$ system has its origin fixed at the (moving) object center of gravity.
-The $z$-axis points upward perpendicularly to the tangent plane of the ellipsoid, and the $x$-axis points towards true north (and not the magnetic north).
-The $y$-axis point towards east.
-
-\vspace{0.1cm}
-
-\item The object-fixed reference frame $\{c\}$ is a moving and rotating coordinate frame that is fixed to the smartphone device. 
-\end{itemize}
-
-The library admit admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure~\ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal} .
-
-\begin{figure}[!h]
-\centering
-\includegraphics[scale=0.55]{images/Schema_situation.png}
-\caption{Definition of the scene with 2 frame:  a fixed frame NED (North East Down) noted $\{a\}$ and a moving frame object noted $\{c\}$}
-\label{Schema_situation}
-\end{figure}
-
-La librairie donne une estimation le quaternion $\mathbi{q}_a^c$
-
 \section{Calibration and initialization}
 
-\section{State of the art}
-
-\subsection{Kinematic and quaternion representation}
-
-A quaternion is a four-dimensional vector, defined as
-
-\begin{center}
-$ \mathbi{q} = \begin{pmatrix} \textbf{q} \\ q_w \end{pmatrix} $ with $ \textbf{q} =  ( q_x, q_y, q_z)^T $. 
-\end{center}
-
-For attitude representation, the quaternion must satisfies a single constraint given by  $|| \mathbi{q} || = 1$.The quaternion kinematics equation in discrete-time is given by :
-
-\begin{equation}
-\mathbi{q}_{k+1} = \Phi_{k+1|k}(\boldsymbol\omega)\mathbi{q}_{k}
-\label{quat_kine}
-\end{equation}
-
-where $\boldsymbol\omega$ is the angular velocity and $\Phi_{k+4|k}$ is the transition matrix from time $k-1$ to $k$ such as:
-
-\begin{equation}
- \Phi_{k+1|k}(\boldsymbol\omega) = \begin{pmatrix}  \cos(0.5||\boldsymbol\omega||\Delta T)I_3 - [\boldsymbol \psi\times] & \boldsymbol\psi \\ -\boldsymbol\psi^T &  \cos(0.5||\boldsymbol\omega||\Delta T)  \end{pmatrix}
-\end{equation}
-
- with,
-
-\begin{equation}
- \boldsymbol \psi = \sin(0.5||\boldsymbol\omega||\Delta T)\frac{\boldsymbol\omega}{||\boldsymbol\omega||}
-\end{equation}
-
-And $[\textbf{v} \times] \in \mathbb{R}^{3\times 3}$ is the skew-symetric matrix of the vector $\textbf{v}$ such as.
-
-\begin{equation}
-[\textbf{v} \times] = \begin{pmatrix} 0 & v_z & -v_y \\ -v_z & 0 & v_x \\ v_y & -v_x & 0 \end{pmatrix} 
-\label{skewsymmat}
-\end{equation}
-Then, the attitude matrix $R \in SO(3) $ related to the quaternion $\mathbi{q}$ is given by :
-
-\begin{equation}
-R(\mathbi{q}) = (q_w^2 - |\textbf{q}|^2)I + 2 \textbf{q}\textbf{q}^T - 2q_w[\textbf{q} \times]
-\label{quat_to_rot}
-\end{equation}
-
-Where $SO(3)$ is the special orthogonal group defined by:
-
-\begin{center}
-$SO(3) = \{ R \in \mathbb{R}^{ 3 \times 3} | RR^T = R^TR = I, det(R) = 1 \}$
-\end{center}
-
-\subsection{Most Common Fusion Methods}
-
-\subsubsection{Kalman Filtering Methods}
-
-Introduced in 1960 by Kalman \cite{kalman_new_1960}, the Kalman filter is an algorithm that uses a series of measurements observed over time, containing noise, in order to estimate unknown variables (state vector).
-%The goal is to obtain more precise results than those based on a single measurement alone.
-Based on a model it is possible to obtain a more robust estimation of the state vector.
-Consider the following linear system, described by the difference equation and the observation model with additive noise:
-
-\begin{equation}
-\left\{ \begin{array}{l}
-\textbf{x}_{k} = F_{k}\textbf{x}_{k-1}+B_{k}\textbf{u}_{k}+G_{k}\textbf{v}_{k}\\
-\textbf{z}_{k} = H_{k}\textbf{x}_{k}+D_{k}\textbf{w}_{k}
-\end{array} \right.
-\label{linear_system}
-\end{equation}
-
-with,
-
-\begin{center}
-\begin{tabular}{rl}
-$\textbf{x}_{k} $ &State vector at time $k$\\
-$\textbf{z}_{k} $ &Measurement vector at time $k$\\
-$\textbf{u}_{k} $ & Input vector (e.i.: from sensor) at time $k$ \\
-$\textbf{v}_{k} $ & State noise vector disturbing the system\\
-$\textbf{w}_{k} $ & Observation noise disturbing the measurement\\
-\end{tabular}
-\end{center}
-
-\vspace{0.2cm}
-
-The figure \ref{State_equation_discrete_case} represents the dynamical system described in \ref{linear_system} .
-A brief presentation of the algorithm is provided here,  see \cite{terejanu2013discrete} for more details.
-The algorithm has 3 important steps :\\
-
-\begin{itemize}
-\item The first step is the \textbf{prediction} of state $\textbf{x}^-_{k}$ at time $k$ and covariance $P^-_{k}$ is given by :
-
-\begin{equation}
-\left\{ \begin{array}{cl}
-\textbf{x}^-_{k} = & F_{k}\textbf{x}^+_{k-1} + B_{k}\textbf{u}_{k} \\
-P^-_{k} = & F_{k}P^+_{k-1}F^T_{k}+G_{k}Q_{k}G^T_{k}
-\end{array}
-\right.
-\end{equation}
-
-\item The \textbf{innovation} is the difference between predicted measurement and real measurement. It measure up the deviation used by the Kalman gain.
-\begin{equation}
-\left\{ \begin{array}{cl}
-\boldsymbol\nu_{k} = & \textbf{z}_{k}-H_{k}\textbf{x}^-_{k} \\
-S_{k} = & R_{k} +H_{k}P^-_{k}H^T_{k}
-\end{array}
-\right.
-\end{equation}
-
-\item And finally, An \textbf{update} is processed to correct the estimate state $\textbf{x}^+_{k} $ and his covariance $P^+_{k}$ with the innovation.
-
-\begin{equation}
-\left\{ \begin{array}{cl}
-\textbf{x}^+_{k} = & \textbf{x}^-_{k} + K_{k}\boldsymbol\nu_{k} \\
-P^+_{k} = &P^-_{k} - K_{k}S_{k}K^T_{k}
-\end{array}
-\right.
-\end{equation}
-
-\end{itemize}
-
-Where $K_k$ is the Kalman gain  representing relative importance of innovation  $\nu_{k}$.
-
-\begin{equation}
-K_{k}=P^-_{k}H^T_{k}S^{-1}
-\end{equation}
-
-%And so on for future state $k+1,...$.
-
-\subsubsection{Nonlinear Observer}
-
-The theory of the state observer was first introduced by Kalman and Bucy for a linear system in a stochastic environment.
-Then Luenberger \cite{david1971introduction} made a general theory of observers for deterministic linear systems, introducing the notions of observer and reduced minimum observer \cite{primbs1996survey}.
-Mahony et al.
-\cite{mahony_nonlinear_2008} have proposed an nonlinear observer called Constant Gain Observer (CGO), termed the explicit complementary filter, that provides attitude estimates as well as gyro bias estimates.
-
-\begin{itemize}
-\item Prediction
-\end{itemize}
-
-First, we propagate the estimate attitude and the gyro bias with equation \ref{prediction_cgo}.
-
-\begin{equation}
-\left\{\begin{array}{l}
-\mathbi{q}_{k}^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro} - \boldsymbol \beta_{k-1}^+)\mathbi{q}_{k-1}^+ \\
-\boldsymbol \beta_k^- = \boldsymbol \beta_{k-1}^+
- \end{array}
-\right.
-\label{prediction_cgo}
-\end{equation}
-
-\begin{itemize}
-\item Correction
-\end{itemize}
-
-The idea is to compute a  constraint  given by a comparison between  an estimate value of a sensor and his real data at a time $k$.
-
-\begin{equation}
-\left\{\begin{array}{l}
-\textbf{v}_{acc}^- = R_(\mathbi{q}_{k}^-)\textbf{a}^a + \boldsymbol  \beta_{acc} \\
-\textbf{v}_{mag}^- = R_(\mathbi{q}_{k}^-)\textbf{m}^a + \boldsymbol  \beta_{mag}
- \end{array}
-\right.
-\end{equation}
-
-All these vectors are normalized.
-
-The '' gain '' is given by,
-
-\begin{equation}
-\boldsymbol \omega_{mes} = -vex(\sum_{i=1}^n \frac{k_i}{2}(\textbf{v}_i(\textbf{v}_i^-)^T -  \textbf{v}_i^-(\textbf{v}_i)^T )) 
-\label{cgo_gain}
-\end{equation}
-
-where the $vex$ operator define by the inverse of the cross product.
-\begin{equation}
-\left\{\begin{array}{l}
-\textbf{a}(\textbf{b})^T -  \textbf{b}(\textbf{a})^T  = [ (\textbf{a} \times \textbf{b}) \times ]\\
-vex([\textbf{a}\times])  = \textbf{a}
- \end{array}
-\right.
-\end{equation}
-
-Then, we the gain is used to correct the predicted attitude $\mathbi{q}_{k}^-$ by the following equation:
-
-\begin{equation}
-\left\{\begin{array}{l}
-\mathbi{q}_{k}^+ = \mathbi{q}_{k}^- + \frac{\Delta T}{2} \Xi(\mathbi{q}_{k}^-) \boldsymbol\omega_{mes}k_p\\
-\boldsymbol \beta_k^+ = \boldsymbol \beta_{k}^-  - \boldsymbol\omega_{mes}k_i
- \end{array}
-\right.
-\end{equation}
-
-One problem of Kalman Filters is to find the initial values in the covariance (matrix $Q$) because it is difficult to relate it to real physical quantities.
-The covariance matrix is not involve in nonlinear observer theory.
-This is one of the reasons there has been an increasing interest for nonlinear observer design the last decades.
-Another reason is their low computational need.
-
-\subsubsection{Methods Based on Solutions of the Wahba's Problem}
-
-The attitude determination problem for vector observation was first formulated as a least squares estimation problem by Wahba \cite{wahba_least_1965} in 1965 :\\
-
-Given the two sets of $n$ vectors $\{ \textbf{r}_{1},...,\textbf{r}_{n} \}$ and $\{ \textbf{b}_{1},...,\textbf{b}_{n} \}$, $n \geqslant 2 $, where each pair $(\textbf{r}_{i},\textbf{b}_{i})$ corresponds to a generalised vector, $\textbf{x}_{i}$, find the proper orthogonal matrix, $A$, which bring the first into the best least squares coincidence with the second.
-That is, find $A$ which minimises the function $J$ defined as:
-
-\begin{equation}
-J(A) = \frac{1}{2} \sum_{i=1}^n a_i||b_i-Ar_i||^2
-\label{Wahba_loss_function}
-\end{equation}
-
-subject to the constraint $A^TA = I$ and $det(A)=1$. The equation (\ref{Wahba_loss_function}) can be written with a quaternion representation which gives:
-
-\begin{equation}
-J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i||b_i-R(\mathbi{q})r_i||^2
-\label{QUEST}
-\end{equation}
-
-The matrix $R(\mathbi{q})$, is the rotation matrix that corresponds to the quaternion $\mathbi{q}$ (equation \ref{quat_to_rot}). The QUEST algorithm are conducted in the following manner. Let
-
-\begin{equation}
-B = \sum_{i=1}^N a_i\textbf{b}_i \textbf{r}_i^T
-\end{equation}
-
-Then, we can rewrite the equation (\ref{QUEST}) as follow
-
-\begin{equation}
-J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i  - tr(R(\mathbi{q})B^T)
-\label{QUEST_bis}
-\end{equation}
-
-Due to the fact that the representation of the attitude matrix is a homogenous quadratic function of $\mathbi{q}$, we can say that
-
-\begin{equation}
-tr(R(\mathbi{q})B^T) = \mathbi{q}^TK\mathbi{q}
-\label{QUEST_ter}
-\end{equation}
-
-where $K$ is the symmetric traceless matrix such as
-
-\begin{equation}
-K = \begin{pmatrix} S-tr(B)I_3 & \textbf{z} \\ \textbf{z}^T & tr(B)
-\end{pmatrix}
-\end{equation}
-
-with
-
-\begin{equation}
-\left\{\begin{array}{l}
-S = B + B^T\\
-\textbf{z} = \sum_ia_i\textbf{b}_i\times\textbf{r}_i
- \end{array}
-\right.
-\end{equation}
-
-The minimisation problem from Equation (\ref{QUEST_bis}) is now a maximisation problem.
-thus, the solution $\mathbi{q}_{opt}$ is the eingen vector correspond to the largest eigen value of $K$ \cite{markley1999estimate}.
-The main problem of QUEST is that it does not take into account of all the previous observations performed since the beginning.
-A recursive QUEST algorithm (REQUEST) proposed by Bar-Itzhack manages this problem.
-According to the REQUEST algorithm, the propagation of $K_k$ is given by
-
-\begin{equation}
-K_k^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro})K_{k-1}^+\Phi_{k|k-1}^T(\boldsymbol\omega_{gyro})
-\end{equation}
-
-We compute $\delta K_k $ as
-
-\begin{equation}
-\delta K_k = \frac{1}{\sum_ia_i}\begin{pmatrix} \delta S_k-\delta\sigma_kI_3 & \delta\textbf{z}_k \\ \delta\textbf{z}_k^T & \delta\sigma_k
-\end{pmatrix}
-\end{equation}
-
-with,
-
-\begin{equation}
-\begin{array}{cc}
-\delta B_k = \sum_i a_i \textbf{b}_i\textbf{r}_i^T & \delta S_k = B_k + B_k^T\\
-\delta \textbf{z}_k = \sum_ia_i\textbf{b}_i\times\textbf{r}_i & \delta \sigma_k = tr(B_k)
-\end{array}
-\end{equation}
-
-Where $\textbf{b}_i$ is the observation vector at time $k$. Then we update the K-matrix.
-
-\begin{equation}
-K_k^+ = \frac{\rho_k m_{k-1}}{m_k}K_k^- + \frac{1}{m_k}\delta K_k
-\end{equation}
-
-A drawback is that the value of $\rho_k$ comes from heuristics.
-Choukroun revisit REQUEST provides an improvement with his Optimal REQUEST \cite{choukroun_novel_2003} where an optimal $\rho_k^*$ is computed which gives better results.
-Here we choose to implement this version of REQUEST.
-Contrary to other methods, the QUEST algorithms do not estimate the gyroscope bias.
-Which can make it less robust in some cases.
-There are also other methods based on Wahba's problem like the Davenport's q-method \cite{weighted1971nasa} proposed in 1971 considered more robust than QUEST.
-Or the method ESOQ proposed by Mortari \cite{mortari1997esoq}  in 1997.
-But in this study we limit our investigation to REQUEST.\\
-
-\subsubsection{Particle Filtering Method}
-
-Particle filtering method is based on Monte-Carlo method\cite{metropolis1949monte}.
-The main idea is to use a large number of samples called particles to estimate a nonlinear function \cite{chen_bayesian_2003} and \cite{terejanu2009tutorial}.
-The set of $N$ particle with associated weight at time $k$ are denoted by $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$.
-Initially, the particle are drawn from a proposal distribution.
-The PF is divided into three step, prediction, update, resampling, that constitute a filter cycle:
-
-\begin{itemize}
-\item Prediction
-\end{itemize}
-
-The first step is to propagate all particles $\{\textbf{x}^{(i)}_{k-1},w^{(i)}_{k-1}\}$ to $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$ with the nonlinear function $f$ while the weight remain unchanged :
-
-\begin{equation}
-\textbf{x}^{(i)}_{k} = f(\textbf{x}^{(i)}_{k-1},\textbf{u}_{k-1}) + \textbf{w}_{k-1}^{(i)}\\
-\end{equation}
-
-where $N$ sample $\textbf{w}_{k-1}^{(i)}$ of the process noise (supposed gaussian in our case) are drawn.
-
-\begin{itemize}
-\item Update/Correction
-\end{itemize}
-
-At the update step, the weights associated with each particle is updated based on the likelihood function $p( \textbf{z}_k| \textbf{x}_k^{(i)})$.
-\begin{equation}
-\left\{ 
-\begin{array}{l}
-w_k^{(i)} = w_{k-1}^{(i)} p( \textbf{z}_k| \textbf{x}_k^{(i)})\\
-\tilde{w}_k^{(i)} = \frac{w_{k}^{(i)}}{\sum_iw_{k}^{(i)}}
-\end{array}
-\right.
-\end{equation}
-
-The weight are normalised such as we have $\sum_i\tilde{w}_k^{(i)}=1$. To finish, the mean is computed with the following equation:
-
-\begin{equation}
- \textbf{x}_k^+ = \sum_{i=1}^N\tilde{w}_k^{(i)} \textbf{x}_k^{(i)}
- \end{equation}
-
-\begin{itemize}
-\item Resampling/Regularisation
-\end{itemize}
-
-A common problem with PF is the degeneracy phenomenon.
-After some iterations, more and more particles will have a weak weight, which make them negligible and useless.
-To overcome this problem, a regularisation step called resampling is processed.
-It is not needed to perform PF but it allows to maintain a good performance relative to degeneracy problem.
-The idea of resampling is to eliminate particles having feeble weight and focus on strong weight particles.
-In order to measure degeneracy, the number of effective samples $N_{eff}$ is computed (equation \ref{N_eff}).
-If  $N_{eff}$ is lower than a threshold, the resampling step is processed.
-
-\begin{equation}
- N_{eff} = \sum_{i=1}^N\frac{1}{(\tilde{w}_k^{(i)})^2}
- \label{N_eff}
- \end{equation}
-
-The choice of the threshold is important.
-Indeed, a high threshold gives $N_{eff} \approx N$, but that does not mean necessarily we have a lot of efficient particles.
-On the contrary, it means that most of the particle are identical making large computations unnecessary.
-Moreover, with a low threshold, many particles are neglected making computations less efficient.
-In general, the threshold is fixed at  $\frac{N}{2}$ or $\frac{N}{3}$.\\
-
-There are several types of resampling, a comparison is made in \cite{douc2005comparison}.
-We took the resampling algorithm described in \cite{arulampalam2002tutorial} because it is the most commun.
-Cheng, Yang and Crassidis have propose an attitute estimation with a particle filter method \cite{cheng_particle_2010}.
-We have also considered the revisited algorithm of Cheng \cite{chang_particle_2014} based on bootstrap filter\cite{gordon1993novel}.
-We have chosen to implement this one because they use the same approach as USQUE \cite{crassidis_unscented_2003} and MEKF\cite{markley2003attitude}.
-We decided to explore the performance of this three algorithms with the same model.
-\\
-
-The particle filter provides an estimate of the nonlinear system without assuming that the noise is Gaussian.
-So it is more robust to any situation, which makes it more performant in the context of a real applications.
-But an important computation time is required and that is a limitation for a real-time application.
-
-\section{Performance Evaluation}
-
-We have access through our designed ground truth to the true attitude $\mathbi{q}^{real}_k$ at time $k$ of the reference frame $\{c\}$ related to $\{a\}$.
-By computation, each method returns an estimate attitude $\mathbi{q}^{est}_k$ at time $k$ (see figure \ref{test_method}).
-
-\begin{equation}
- \delta \mathbi{q}_k = \mathbi{q}^{real}_k \otimes (\mathbi{q}^{est}_k)^{-1} =  \begin{pmatrix}  \delta \textbf{q}_{k} \\ \delta q_{w_k}  \end{pmatrix}
-\end{equation}
-
-The error-MRP vector related to $ \delta \mathbi{q}_k$ is given by
-
-\begin{equation}
- \delta \textbf{p}_k  = 4[\frac{\delta \textbf{q}_{k}  }{(1+\delta q_{w_k})}]
-\end{equation}
-
-Thus, we can calculate the attitude error given by the root mean square (RMS) of $\delta \textbf{p}_k$:
-
-\begin{equation}
-\epsilon_k  = RMS (\delta \textbf{p}_k) = \sqrt{\frac{1}{3}( \delta p_{k_x}^2 +\delta p_{k_y}^2+\delta p_{k_z}^2)  }
-\label{error_definition}
-\end{equation}
-
-The mean of the error is given by :
-
-\begin{equation}
-\bar{\epsilon}  = \frac{1}{n}\sum_{i=1}^n \epsilon_i \\
-\label{mean_var_error_def}
-\end{equation}
-
-With $n$ the number of data at our disposal. We use this information (mean error) for our performance study.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[scale=0.40]{images/test_method.png}
-\caption{Algorithm designed to compute attitude estimation methods' performances}
-\label{test_method}
-\end{figure}
-
-To differentiate the methods, we assigned  some ''scores'' in order to highlight the behaviour of these methods in different configurations. We defined these ''scores'' as:
-
-\begin{align}
-s_1 = \bar{\epsilon}\\
-s_2 = \frac{1}{\bar{\epsilon} \times \bar{\tau}}
-\label{score}
-\end{align}
-
-With $\bar{\epsilon}$ average attitude error (voir \ref{error_definition}) et $\bar{\tau}$ the average computation time. \\
-
-\begin{figure}[!h]
-
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.85\linewidth]{images/histo_s1_add.png}
-  \caption{Mean Attitude Error (score $s_1$)}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.85\linewidth]{images/histo_s2_add.png}
-  \caption{Ratio Error times Computation Time (score $s_2$)}
-
-\end{subfigure}
-\caption{Scores of methods where noise is AWGN with 3 strength}
-
-\end{figure}
-
-\section{Conclusion}
-
-notre library est trop bien...
-
-\bibliographystyle{ieeetr}
-
-\end{document}
+A toi de jouer Vivien...

### 2016-05-09 Article/conclusion.tex
adding different file for article

\section{Conclusion}

notre library est trop bien...

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
    Text for this section \ldots

\section*{Acknowledgements}
This work was supported by the ARAP program of ASTAR and by the university Sorbonne UPMC.

\bibliographystyle{bmc-mathphys} 
\bibliography{bmc_article}

\end{backmatter}

### 2016-05-09 Article/introduction.tex
adding different file for article

--- 2016-05-09 Article/main.tex
@@
-\documentclass[twocolumn]{bmcart}
-
-\usepackage[utf8]{inputenc} %unicode support
-
-\usepackage{graphics} % for pdf, bitmapped graphics files
-\usepackage{epsfig} % for postscript graphics files
-%\usepackage{mathptmx} % assumes new font selection scheme installed
-%\usepackage{times} % assumes new font selection scheme installed
-\usepackage{amsmath} % assumes amsmath package installed
-\usepackage{amssymb}  % assumes amsmath package installed
-\usepackage{subcaption}
-
-\startlocaldefs
-
-\def\mathbi#1{\textbf{\em #1}}
-\makeatletter
-\newcommand*{\rom}[1]{\expandafter\@slowromancap\romannumeral #1@}
-\makeatother
-
-\endlocaldefs
-
-%%% Begin ...
-\begin{document}
-
-%%% Start of article front matter
-\begin{frontmatter}
-
-\begin{fmbox}
-\dochead{Research}
-
-\author[
-   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
-   corref={aff1},     
-   noteref={n1},                        % id's of article notes, if any
-   email={nizar.ouarti@ipal.cnrs.fr}   % email address
-]{\inits{NO}\fnm{Nizar} \snm{Ouarti}}
-\author[
-   addressref={aff1},   corref={aff1},  
-   email={braud.thomas@ipal.cnrs.fr}
-]{\inits{TB}\fnm{Thomas} \snm{Braud}}
-\author[
-   addressref={aff1},   corref={aff1},  
-   email={vivien.billaud@ipal.cnrs.fr}
-]{\inits{VB}\fnm{Vivien} \snm{Billaud}}
-
-\address[id=aff1]{%                           % unique id
-  \orgname{IPAL (Sorbonne UPMC, CNRS, Astar, NUS, UJF, IMT)}, % university, etc
-  \city{Singapore}
-  }
-
-\end{fmbox}% comment this for two column layout
-
- \title{OpenMotion: An open source library for attitude estimation.}
-
-\maketitle
-
 \begin{abstractbox}
@@
 \end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The keywords begin here                  %%
-%%                                          %%
-%% Put each keyword in separate \kwd{}.     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \section{Introduction}
-
-\section{Global Presentation of the Library}
-
-\texttt{OpenMotion} est une librairie open source qui permet a l'utilisateur d'estimer l'orientation du telephone a partir des donnee de l'IMU.
-La librairie ne supporte pas encore les translations, nous nous sommes concentre sur l'attitude.
-
-\begin{itemize}
-\item The North East Down (NED) frame $\{a\}$ system has its origin fixed at the (moving) object center of gravity.
-The $z$-axis points upward perpendicularly to the tangent plane of the ellipsoid, and the $x$-axis points towards true north (and not the magnetic north).
-The $y$-axis point towards east.
-
-\vspace{0.1cm}
-
-\item The object-fixed reference frame $\{c\}$ is a moving and rotating coordinate frame that is fixed to the smartphone device. 
-\end{itemize}
-
-The library admit admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure~\ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal} .
-
-\begin{figure}[!h]
-\centering
-\includegraphics[scale=0.55]{images/Schema_situation.png}
-\caption{Definition of the scene with 2 frame:  a fixed frame NED (North East Down) noted $\{a\}$ and a moving frame object noted $\{c\}$}
-\label{Schema_situation}
-\end{figure}
-
-La librairie donne une estimation le quaternion $\mathbi{q}_a^c$
-
-\section{Calibration and initialization}
-
-\section{State of the art}
-
-\subsection{Kinematic and quaternion representation}
-
-A quaternion is a four-dimensional vector, defined as
-
-\begin{center}
-$ \mathbi{q} = \begin{pmatrix} \textbf{q} \\ q_w \end{pmatrix} $ with $ \textbf{q} =  ( q_x, q_y, q_z)^T $. 
-\end{center}
-
-For attitude representation, the quaternion must satisfies a single constraint given by  $|| \mathbi{q} || = 1$.The quaternion kinematics equation in discrete-time is given by :
-
-\begin{equation}
-\mathbi{q}_{k+1} = \Phi_{k+1|k}(\boldsymbol\omega)\mathbi{q}_{k}
-\label{quat_kine}
-\end{equation}
-
-where $\boldsymbol\omega$ is the angular velocity and $\Phi_{k+4|k}$ is the transition matrix from time $k-1$ to $k$ such as:
-
-\begin{equation}
- \Phi_{k+1|k}(\boldsymbol\omega) = \begin{pmatrix}  \cos(0.5||\boldsymbol\omega||\Delta T)I_3 - [\boldsymbol \psi\times] & \boldsymbol\psi \\ -\boldsymbol\psi^T &  \cos(0.5||\boldsymbol\omega||\Delta T)  \end{pmatrix}
-\end{equation}
-
- with,
-
-\begin{equation}
- \boldsymbol \psi = \sin(0.5||\boldsymbol\omega||\Delta T)\frac{\boldsymbol\omega}{||\boldsymbol\omega||}
-\end{equation}
-
-And $[\textbf{v} \times] \in \mathbb{R}^{3\times 3}$ is the skew-symetric matrix of the vector $\textbf{v}$ such as.
-
-\begin{equation}
-[\textbf{v} \times] = \begin{pmatrix} 0 & v_z & -v_y \\ -v_z & 0 & v_x \\ v_y & -v_x & 0 \end{pmatrix} 
-\label{skewsymmat}
-\end{equation}
-Then, the attitude matrix $R \in SO(3) $ related to the quaternion $\mathbi{q}$ is given by :
-
-\begin{equation}
-R(\mathbi{q}) = (q_w^2 - |\textbf{q}|^2)I + 2 \textbf{q}\textbf{q}^T - 2q_w[\textbf{q} \times]
-\label{quat_to_rot}
-\end{equation}
-
-Where $SO(3)$ is the special orthogonal group defined by:
-
-\begin{center}
-$SO(3) = \{ R \in \mathbb{R}^{ 3 \times 3} | RR^T = R^TR = I, det(R) = 1 \}$
-\end{center}
-
-\subsection{Most Common Fusion Methods}
-
-\subsubsection{Kalman Filtering Methods}
-
-Introduced in 1960 by Kalman \cite{kalman_new_1960}, the Kalman filter is an algorithm that uses a series of measurements observed over time, containing noise, in order to estimate unknown variables (state vector).
-%The goal is to obtain more precise results than those based on a single measurement alone.
-Based on a model it is possible to obtain a more robust estimation of the state vector.
-Consider the following linear system, described by the difference equation and the observation model with additive noise:
-
-\begin{equation}
-\left\{ \begin{array}{l}
-\textbf{x}_{k} = F_{k}\textbf{x}_{k-1}+B_{k}\textbf{u}_{k}+G_{k}\textbf{v}_{k}\\
-\textbf{z}_{k} = H_{k}\textbf{x}_{k}+D_{k}\textbf{w}_{k}
-\end{array} \right.
-\label{linear_system}
-\end{equation}
-
-with,
-
-\begin{center}
-\begin{tabular}{rl}
-$\textbf{x}_{k} $ &State vector at time $k$\\
-$\textbf{z}_{k} $ &Measurement vector at time $k$\\
-$\textbf{u}_{k} $ & Input vector (e.i.: from sensor) at time $k$ \\
-$\textbf{v}_{k} $ & State noise vector disturbing the system\\
-$\textbf{w}_{k} $ & Observation noise disturbing the measurement\\
-\end{tabular}
-\end{center}
-
-\vspace{0.2cm}
-
-The figure \ref{State_equation_discrete_case} represents the dynamical system described in \ref{linear_system} .
-A brief presentation of the algorithm is provided here,  see \cite{terejanu2013discrete} for more details.
-The algorithm has 3 important steps :\\
-
-\begin{itemize}
-\item The first step is the \textbf{prediction} of state $\textbf{x}^-_{k}$ at time $k$ and covariance $P^-_{k}$ is given by :
-
-\begin{equation}
-\left\{ \begin{array}{cl}
-\textbf{x}^-_{k} = & F_{k}\textbf{x}^+_{k-1} + B_{k}\textbf{u}_{k} \\
-P^-_{k} = & F_{k}P^+_{k-1}F^T_{k}+G_{k}Q_{k}G^T_{k}
-\end{array}
-\right.
-\end{equation}
-
-\item The \textbf{innovation} is the difference between predicted measurement and real measurement. It measure up the deviation used by the Kalman gain.
-\begin{equation}
-\left\{ \begin{array}{cl}
-\boldsymbol\nu_{k} = & \textbf{z}_{k}-H_{k}\textbf{x}^-_{k} \\
-S_{k} = & R_{k} +H_{k}P^-_{k}H^T_{k}
-\end{array}
-\right.
-\end{equation}
-
-\item And finally, An \textbf{update} is processed to correct the estimate state $\textbf{x}^+_{k} $ and his covariance $P^+_{k}$ with the innovation.
-
-\begin{equation}
-\left\{ \begin{array}{cl}
-\textbf{x}^+_{k} = & \textbf{x}^-_{k} + K_{k}\boldsymbol\nu_{k} \\
-P^+_{k} = &P^-_{k} - K_{k}S_{k}K^T_{k}
-\end{array}
-\right.
-\end{equation}
-
-\end{itemize}
-
-Where $K_k$ is the Kalman gain  representing relative importance of innovation  $\nu_{k}$.
-
-\begin{equation}
-K_{k}=P^-_{k}H^T_{k}S^{-1}
-\end{equation}
-
-%And so on for future state $k+1,...$.
-
-\subsubsection{Nonlinear Observer}
-
-The theory of the state observer was first introduced by Kalman and Bucy for a linear system in a stochastic environment.
-Then Luenberger \cite{david1971introduction} made a general theory of observers for deterministic linear systems, introducing the notions of observer and reduced minimum observer \cite{primbs1996survey}.
-Mahony et al.
-\cite{mahony_nonlinear_2008} have proposed an nonlinear observer called Constant Gain Observer (CGO), termed the explicit complementary filter, that provides attitude estimates as well as gyro bias estimates.
-
-\begin{itemize}
-\item Prediction
-\end{itemize}
-
-First, we propagate the estimate attitude and the gyro bias with equation \ref{prediction_cgo}.
-
-\begin{equation}
-\left\{\begin{array}{l}
-\mathbi{q}_{k}^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro} - \boldsymbol \beta_{k-1}^+)\mathbi{q}_{k-1}^+ \\
-\boldsymbol \beta_k^- = \boldsymbol \beta_{k-1}^+
- \end{array}
-\right.
-\label{prediction_cgo}
-\end{equation}
-
-\begin{itemize}
-\item Correction
-\end{itemize}
-
-The idea is to compute a  constraint  given by a comparison between  an estimate value of a sensor and his real data at a time $k$.
-
-\begin{equation}
-\left\{\begin{array}{l}
-\textbf{v}_{acc}^- = R_(\mathbi{q}_{k}^-)\textbf{a}^a + \boldsymbol  \beta_{acc} \\
-\textbf{v}_{mag}^- = R_(\mathbi{q}_{k}^-)\textbf{m}^a + \boldsymbol  \beta_{mag}
- \end{array}
-\right.
-\end{equation}
-
-All these vectors are normalized.
-
-The '' gain '' is given by,
-
-\begin{equation}
-\boldsymbol \omega_{mes} = -vex(\sum_{i=1}^n \frac{k_i}{2}(\textbf{v}_i(\textbf{v}_i^-)^T -  \textbf{v}_i^-(\textbf{v}_i)^T )) 
-\label{cgo_gain}
-\end{equation}
-
-where the $vex$ operator define by the inverse of the cross product.
-\begin{equation}
-\left\{\begin{array}{l}
-\textbf{a}(\textbf{b})^T -  \textbf{b}(\textbf{a})^T  = [ (\textbf{a} \times \textbf{b}) \times ]\\
-vex([\textbf{a}\times])  = \textbf{a}
- \end{array}
-\right.
-\end{equation}
-
-Then, we the gain is used to correct the predicted attitude $\mathbi{q}_{k}^-$ by the following equation:
-
-\begin{equation}
-\left\{\begin{array}{l}
-\mathbi{q}_{k}^+ = \mathbi{q}_{k}^- + \frac{\Delta T}{2} \Xi(\mathbi{q}_{k}^-) \boldsymbol\omega_{mes}k_p\\
-\boldsymbol \beta_k^+ = \boldsymbol \beta_{k}^-  - \boldsymbol\omega_{mes}k_i
- \end{array}
-\right.
-\end{equation}
-
-One problem of Kalman Filters is to find the initial values in the covariance (matrix $Q$) because it is difficult to relate it to real physical quantities.
-The covariance matrix is not involve in nonlinear observer theory.
-This is one of the reasons there has been an increasing interest for nonlinear observer design the last decades.
-Another reason is their low computational need.
-
-\subsubsection{Methods Based on Solutions of the Wahba's Problem}
-
-The attitude determination problem for vector observation was first formulated as a least squares estimation problem by Wahba \cite{wahba_least_1965} in 1965 :\\
-
-Given the two sets of $n$ vectors $\{ \textbf{r}_{1},...,\textbf{r}_{n} \}$ and $\{ \textbf{b}_{1},...,\textbf{b}_{n} \}$, $n \geqslant 2 $, where each pair $(\textbf{r}_{i},\textbf{b}_{i})$ corresponds to a generalised vector, $\textbf{x}_{i}$, find the proper orthogonal matrix, $A$, which bring the first into the best least squares coincidence with the second.
-That is, find $A$ which minimises the function $J$ defined as:
-
-\begin{equation}
-J(A) = \frac{1}{2} \sum_{i=1}^n a_i||b_i-Ar_i||^2
-\label{Wahba_loss_function}
-\end{equation}
-
-subject to the constraint $A^TA = I$ and $det(A)=1$. The equation (\ref{Wahba_loss_function}) can be written with a quaternion representation which gives:
-
-\begin{equation}
-J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i||b_i-R(\mathbi{q})r_i||^2
-\label{QUEST}
-\end{equation}
-
-The matrix $R(\mathbi{q})$, is the rotation matrix that corresponds to the quaternion $\mathbi{q}$ (equation \ref{quat_to_rot}). The QUEST algorithm are conducted in the following manner. Let
-
-\begin{equation}
-B = \sum_{i=1}^N a_i\textbf{b}_i \textbf{r}_i^T
-\end{equation}
-
-Then, we can rewrite the equation (\ref{QUEST}) as follow
-
-\begin{equation}
-J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i  - tr(R(\mathbi{q})B^T)
-\label{QUEST_bis}
-\end{equation}
-
-Due to the fact that the representation of the attitude matrix is a homogenous quadratic function of $\mathbi{q}$, we can say that
-
-\begin{equation}
-tr(R(\mathbi{q})B^T) = \mathbi{q}^TK\mathbi{q}
-\label{QUEST_ter}
-\end{equation}
-
-where $K$ is the symmetric traceless matrix such as
-
-\begin{equation}
-K = \begin{pmatrix} S-tr(B)I_3 & \textbf{z} \\ \textbf{z}^T & tr(B)
-\end{pmatrix}
-\end{equation}
-
-with
-
-\begin{equation}
-\left\{\begin{array}{l}
-S = B + B^T\\
-\textbf{z} = \sum_ia_i\textbf{b}_i\times\textbf{r}_i
- \end{array}
-\right.
-\end{equation}
-
-The minimisation problem from Equation (\ref{QUEST_bis}) is now a maximisation problem.
-thus, the solution $\mathbi{q}_{opt}$ is the eingen vector correspond to the largest eigen value of $K$ \cite{markley1999estimate}.
-The main problem of QUEST is that it does not take into account of all the previous observations performed since the beginning.
-A recursive QUEST algorithm (REQUEST) proposed by Bar-Itzhack manages this problem.
-According to the REQUEST algorithm, the propagation of $K_k$ is given by
-
-\begin{equation}
-K_k^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro})K_{k-1}^+\Phi_{k|k-1}^T(\boldsymbol\omega_{gyro})
-\end{equation}
-
-We compute $\delta K_k $ as
-
-\begin{equation}
-\delta K_k = \frac{1}{\sum_ia_i}\begin{pmatrix} \delta S_k-\delta\sigma_kI_3 & \delta\textbf{z}_k \\ \delta\textbf{z}_k^T & \delta\sigma_k
-\end{pmatrix}
-\end{equation}
-
-with,
-
-\begin{equation}
-\begin{array}{cc}
-\delta B_k = \sum_i a_i \textbf{b}_i\textbf{r}_i^T & \delta S_k = B_k + B_k^T\\
-\delta \textbf{z}_k = \sum_ia_i\textbf{b}_i\times\textbf{r}_i & \delta \sigma_k = tr(B_k)
-\end{array}
-\end{equation}
-
-Where $\textbf{b}_i$ is the observation vector at time $k$. Then we update the K-matrix.
-
-\begin{equation}
-K_k^+ = \frac{\rho_k m_{k-1}}{m_k}K_k^- + \frac{1}{m_k}\delta K_k
-\end{equation}
-
-A drawback is that the value of $\rho_k$ comes from heuristics.
-Choukroun revisit REQUEST provides an improvement with his Optimal REQUEST \cite{choukroun_novel_2003} where an optimal $\rho_k^*$ is computed which gives better results.
-Here we choose to implement this version of REQUEST.
-Contrary to other methods, the QUEST algorithms do not estimate the gyroscope bias.
-Which can make it less robust in some cases.
-There are also other methods based on Wahba's problem like the Davenport's q-method \cite{weighted1971nasa} proposed in 1971 considered more robust than QUEST.
-Or the method ESOQ proposed by Mortari \cite{mortari1997esoq}  in 1997.
-But in this study we limit our investigation to REQUEST.\\
-
-\subsubsection{Particle Filtering Method}
-
-Particle filtering method is based on Monte-Carlo method\cite{metropolis1949monte}.
-The main idea is to use a large number of samples called particles to estimate a nonlinear function \cite{chen_bayesian_2003} and \cite{terejanu2009tutorial}.
-The set of $N$ particle with associated weight at time $k$ are denoted by $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$.
-Initially, the particle are drawn from a proposal distribution.
-The PF is divided into three step, prediction, update, resampling, that constitute a filter cycle:
-
-\begin{itemize}
-\item Prediction
-\end{itemize}
-
-The first step is to propagate all particles $\{\textbf{x}^{(i)}_{k-1},w^{(i)}_{k-1}\}$ to $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$ with the nonlinear function $f$ while the weight remain unchanged :
-
-\begin{equation}
-\textbf{x}^{(i)}_{k} = f(\textbf{x}^{(i)}_{k-1},\textbf{u}_{k-1}) + \textbf{w}_{k-1}^{(i)}\\
-\end{equation}
-
-where $N$ sample $\textbf{w}_{k-1}^{(i)}$ of the process noise (supposed gaussian in our case) are drawn.
-
-\begin{itemize}
-\item Update/Correction
-\end{itemize}
-
-At the update step, the weights associated with each particle is updated based on the likelihood function $p( \textbf{z}_k| \textbf{x}_k^{(i)})$.
-\begin{equation}
-\left\{ 
-\begin{array}{l}
-w_k^{(i)} = w_{k-1}^{(i)} p( \textbf{z}_k| \textbf{x}_k^{(i)})\\
-\tilde{w}_k^{(i)} = \frac{w_{k}^{(i)}}{\sum_iw_{k}^{(i)}}
-\end{array}
-\right.
-\end{equation}
-
-The weight are normalised such as we have $\sum_i\tilde{w}_k^{(i)}=1$. To finish, the mean is computed with the following equation:
-
-\begin{equation}
- \textbf{x}_k^+ = \sum_{i=1}^N\tilde{w}_k^{(i)} \textbf{x}_k^{(i)}
- \end{equation}
-
-\begin{itemize}
-\item Resampling/Regularisation
-\end{itemize}
-
-A common problem with PF is the degeneracy phenomenon.
-After some iterations, more and more particles will have a weak weight, which make them negligible and useless.
-To overcome this problem, a regularisation step called resampling is processed.
-It is not needed to perform PF but it allows to maintain a good performance relative to degeneracy problem.
-The idea of resampling is to eliminate particles having feeble weight and focus on strong weight particles.
-In order to measure degeneracy, the number of effective samples $N_{eff}$ is computed (equation \ref{N_eff}).
-If  $N_{eff}$ is lower than a threshold, the resampling step is processed.
-
-\begin{equation}
- N_{eff} = \sum_{i=1}^N\frac{1}{(\tilde{w}_k^{(i)})^2}
- \label{N_eff}
- \end{equation}
-
-The choice of the threshold is important.
-Indeed, a high threshold gives $N_{eff} \approx N$, but that does not mean necessarily we have a lot of efficient particles.
-On the contrary, it means that most of the particle are identical making large computations unnecessary.
-Moreover, with a low threshold, many particles are neglected making computations less efficient.
-In general, the threshold is fixed at  $\frac{N}{2}$ or $\frac{N}{3}$.\\
-
-There are several types of resampling, a comparison is made in \cite{douc2005comparison}.
-We took the resampling algorithm described in \cite{arulampalam2002tutorial} because it is the most commun.
-Cheng, Yang and Crassidis have propose an attitute estimation with a particle filter method \cite{cheng_particle_2010}.
-We have also considered the revisited algorithm of Cheng \cite{chang_particle_2014} based on bootstrap filter\cite{gordon1993novel}.
-We have chosen to implement this one because they use the same approach as USQUE \cite{crassidis_unscented_2003} and MEKF\cite{markley2003attitude}.
-We decided to explore the performance of this three algorithms with the same model.
-\\
-
-The particle filter provides an estimate of the nonlinear system without assuming that the noise is Gaussian.
-So it is more robust to any situation, which makes it more performant in the context of a real applications.
-But an important computation time is required and that is a limitation for a real-time application.
-
-\section{Performance Evaluation}
-
-We have access through our designed ground truth to the true attitude $\mathbi{q}^{real}_k$ at time $k$ of the reference frame $\{c\}$ related to $\{a\}$.
-By computation, each method returns an estimate attitude $\mathbi{q}^{est}_k$ at time $k$ (see figure \ref{test_method}).
-
-\begin{equation}
- \delta \mathbi{q}_k = \mathbi{q}^{real}_k \otimes (\mathbi{q}^{est}_k)^{-1} =  \begin{pmatrix}  \delta \textbf{q}_{k} \\ \delta q_{w_k}  \end{pmatrix}
-\end{equation}
-
-The error-MRP vector related to $ \delta \mathbi{q}_k$ is given by
-
-\begin{equation}
- \delta \textbf{p}_k  = 4[\frac{\delta \textbf{q}_{k}  }{(1+\delta q_{w_k})}]
-\end{equation}
-
-Thus, we can calculate the attitude error given by the root mean square (RMS) of $\delta \textbf{p}_k$:
-
-\begin{equation}
-\epsilon_k  = RMS (\delta \textbf{p}_k) = \sqrt{\frac{1}{3}( \delta p_{k_x}^2 +\delta p_{k_y}^2+\delta p_{k_z}^2)  }
-\label{error_definition}
-\end{equation}
-
-The mean of the error is given by :
-
-\begin{equation}
-\bar{\epsilon}  = \frac{1}{n}\sum_{i=1}^n \epsilon_i \\
-\label{mean_var_error_def}
-\end{equation}
-
-With $n$ the number of data at our disposal. We use this information (mean error) for our performance study.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[scale=0.40]{images/test_method.png}
-\caption{Algorithm designed to compute attitude estimation methods' performances}
-\label{test_method}
-\end{figure}
-
-To differentiate the methods, we assigned  some ''scores'' in order to highlight the behaviour of these methods in different configurations. We defined these ''scores'' as:
-
-\begin{align}
-s_1 = \bar{\epsilon}\\
-s_2 = \frac{1}{\bar{\epsilon} \times \bar{\tau}}
-\label{score}
-\end{align}
-
-With $\bar{\epsilon}$ average attitude error (voir \ref{error_definition}) et $\bar{\tau}$ the average computation time. \\
-
-\begin{figure}[!h]
-
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.85\linewidth]{images/histo_s1_add.png}
-  \caption{Mean Attitude Error (score $s_1$)}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.85\linewidth]{images/histo_s2_add.png}
-  \caption{Ratio Error times Computation Time (score $s_2$)}
-
-\end{subfigure}
-\caption{Scores of methods where noise is AWGN with 3 strength}
-
-\end{figure}
-
-\section{Conclusion}
-
-notre library est trop bien...
-
-\bibliographystyle{ieeetr}
-
-\end{document}

### 2016-05-09 Article/main.tex
adding different file for article

@@
 \documentclass[twocolumn]{bmcart}
+
+%%%%%%%%%%%%
+% Packages
+%%%%%%%%%%%%
 
@@
 \usepackage{epsfig} % for postscript graphics files
-%\usepackage{mathptmx} % assumes new font selection scheme installed
-%\usepackage{times} % assumes new font selection scheme installed
 \usepackage{amsmath} % assumes amsmath package installed
@@
 
+%%%%%%%%%%%%
+% Local definition
+%%%%%%%%%%%%
+
 \startlocaldefs
 
+% quaternion style
 \def\mathbi#1{\textbf{\em #1}}
+
+% roman letter
 \makeatletter
@@
 
-%%% Begin ...
+%%%%%%%%%%%%
+% Begin Document
+%%%%%%%%%%%%
 \begin{document}
 
-%%% Start of article front matter
+%%%%%%%%%%%%
+% Title
+%%%%%%%%%%%%
+
 \begin{frontmatter}
@@
 
-\begin{abstractbox}
+%%%%%%%%%%%%
+% Abstract + Introduction
+%%%%%%%%%%%%
 
-\begin{abstract} % abstract
+\input{introduction.tex}
 
-The abstract goes here
-\end{abstract}
+%%%%%%%%%%%%
+% General Presentation of the library
+%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The keywords begin here                  %%
-%%                                          %%
-%% Put each keyword in separate \kwd{}.     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\input{prez_library.tex}
 
-\begin{keyword}
-\kwd{IMU}
-\kwd{Sensor Fusion}
-\kwd{Sensor Calibration}
-\end{keyword}
+%%%%%%%%%%%%
+% Presentation of the calibration process
+%%%%%%%%%%%%
 
-\end{abstractbox}
+\input{calibration.tex}
 
-\end{frontmatter}
+%%%%%%%%%%%%
+% State of the art
+%%%%%%%%%%%%
 
-\section{Introduction}
+\input{state_of_the_art.tex}
 
-\section{Global Presentation of the Library}
+%%%%%%%%%%%%
+% Conclusion - Discussion
+%%%%%%%%%%%%
 
-\texttt{OpenMotion} est une librairie open source qui permet a l'utilisateur d'estimer l'orientation du telephone a partir des donnee de l'IMU.
-La librairie ne supporte pas encore les translations, nous nous sommes concentre sur l'attitude.
-
-\begin{itemize}
-\item The North East Down (NED) frame $\{a\}$ system has its origin fixed at the (moving) object center of gravity.
-The $z$-axis points upward perpendicularly to the tangent plane of the ellipsoid, and the $x$-axis points towards true north (and not the magnetic north).
-The $y$-axis point towards east.
-
-\vspace{0.1cm}
-
-\item The object-fixed reference frame $\{c\}$ is a moving and rotating coordinate frame that is fixed to the smartphone device. 
-\end{itemize}
-
-The library admit admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure~\ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal} .
-
-\begin{figure}[!h]
-\centering
-\includegraphics[scale=0.55]{images/Schema_situation.png}
-\caption{Definition of the scene with 2 frame:  a fixed frame NED (North East Down) noted $\{a\}$ and a moving frame object noted $\{c\}$}
-\label{Schema_situation}
-\end{figure}
-
-La librairie donne une estimation le quaternion $\mathbi{q}_a^c$
-
-\section{Calibration and initialization}
-
-\section{State of the art}
-
-\subsection{Kinematic and quaternion representation}
-
-A quaternion is a four-dimensional vector, defined as
-
-\begin{center}
-$ \mathbi{q} = \begin{pmatrix} \textbf{q} \\ q_w \end{pmatrix} $ with $ \textbf{q} =  ( q_x, q_y, q_z)^T $. 
-\end{center}
-
-For attitude representation, the quaternion must satisfies a single constraint given by  $|| \mathbi{q} || = 1$.The quaternion kinematics equation in discrete-time is given by :
-
-\begin{equation}
-\mathbi{q}_{k+1} = \Phi_{k+1|k}(\boldsymbol\omega)\mathbi{q}_{k}
-\label{quat_kine}
-\end{equation}
-
-where $\boldsymbol\omega$ is the angular velocity and $\Phi_{k+4|k}$ is the transition matrix from time $k-1$ to $k$ such as:
-
-\begin{equation}
- \Phi_{k+1|k}(\boldsymbol\omega) = \begin{pmatrix}  \cos(0.5||\boldsymbol\omega||\Delta T)I_3 - [\boldsymbol \psi\times] & \boldsymbol\psi \\ -\boldsymbol\psi^T &  \cos(0.5||\boldsymbol\omega||\Delta T)  \end{pmatrix}
-\end{equation}
-
- with,
-
-\begin{equation}
- \boldsymbol \psi = \sin(0.5||\boldsymbol\omega||\Delta T)\frac{\boldsymbol\omega}{||\boldsymbol\omega||}
-\end{equation}
-
-And $[\textbf{v} \times] \in \mathbb{R}^{3\times 3}$ is the skew-symetric matrix of the vector $\textbf{v}$ such as.
-
-\begin{equation}
-[\textbf{v} \times] = \begin{pmatrix} 0 & v_z & -v_y \\ -v_z & 0 & v_x \\ v_y & -v_x & 0 \end{pmatrix} 
-\label{skewsymmat}
-\end{equation}
-Then, the attitude matrix $R \in SO(3) $ related to the quaternion $\mathbi{q}$ is given by :
-
-\begin{equation}
-R(\mathbi{q}) = (q_w^2 - |\textbf{q}|^2)I + 2 \textbf{q}\textbf{q}^T - 2q_w[\textbf{q} \times]
-\label{quat_to_rot}
-\end{equation}
-
-Where $SO(3)$ is the special orthogonal group defined by:
-
-\begin{center}
-$SO(3) = \{ R \in \mathbb{R}^{ 3 \times 3} | RR^T = R^TR = I, det(R) = 1 \}$
-\end{center}
-
-\subsection{Most Common Fusion Methods}
-
-\subsubsection{Kalman Filtering Methods}
-
-Introduced in 1960 by Kalman \cite{kalman_new_1960}, the Kalman filter is an algorithm that uses a series of measurements observed over time, containing noise, in order to estimate unknown variables (state vector).
-%The goal is to obtain more precise results than those based on a single measurement alone.
-Based on a model it is possible to obtain a more robust estimation of the state vector.
-Consider the following linear system, described by the difference equation and the observation model with additive noise:
-
-\begin{equation}
-\left\{ \begin{array}{l}
-\textbf{x}_{k} = F_{k}\textbf{x}_{k-1}+B_{k}\textbf{u}_{k}+G_{k}\textbf{v}_{k}\\
-\textbf{z}_{k} = H_{k}\textbf{x}_{k}+D_{k}\textbf{w}_{k}
-\end{array} \right.
-\label{linear_system}
-\end{equation}
-
-with,
-
-\begin{center}
-\begin{tabular}{rl}
-$\textbf{x}_{k} $ &State vector at time $k$\\
-$\textbf{z}_{k} $ &Measurement vector at time $k$\\
-$\textbf{u}_{k} $ & Input vector (e.i.: from sensor) at time $k$ \\
-$\textbf{v}_{k} $ & State noise vector disturbing the system\\
-$\textbf{w}_{k} $ & Observation noise disturbing the measurement\\
-\end{tabular}
-\end{center}
-
-\vspace{0.2cm}
-
-The figure \ref{State_equation_discrete_case} represents the dynamical system described in \ref{linear_system} .
-A brief presentation of the algorithm is provided here,  see \cite{terejanu2013discrete} for more details.
-The algorithm has 3 important steps :\\
-
-\begin{itemize}
-\item The first step is the \textbf{prediction} of state $\textbf{x}^-_{k}$ at time $k$ and covariance $P^-_{k}$ is given by :
-
-\begin{equation}
-\left\{ \begin{array}{cl}
-\textbf{x}^-_{k} = & F_{k}\textbf{x}^+_{k-1} + B_{k}\textbf{u}_{k} \\
-P^-_{k} = & F_{k}P^+_{k-1}F^T_{k}+G_{k}Q_{k}G^T_{k}
-\end{array}
-\right.
-\end{equation}
-
-\item The \textbf{innovation} is the difference between predicted measurement and real measurement. It measure up the deviation used by the Kalman gain.
-\begin{equation}
-\left\{ \begin{array}{cl}
-\boldsymbol\nu_{k} = & \textbf{z}_{k}-H_{k}\textbf{x}^-_{k} \\
-S_{k} = & R_{k} +H_{k}P^-_{k}H^T_{k}
-\end{array}
-\right.
-\end{equation}
-
-\item And finally, An \textbf{update} is processed to correct the estimate state $\textbf{x}^+_{k} $ and his covariance $P^+_{k}$ with the innovation.
-
-\begin{equation}
-\left\{ \begin{array}{cl}
-\textbf{x}^+_{k} = & \textbf{x}^-_{k} + K_{k}\boldsymbol\nu_{k} \\
-P^+_{k} = &P^-_{k} - K_{k}S_{k}K^T_{k}
-\end{array}
-\right.
-\end{equation}
-
-\end{itemize}
-
-Where $K_k$ is the Kalman gain  representing relative importance of innovation  $\nu_{k}$.
-
-\begin{equation}
-K_{k}=P^-_{k}H^T_{k}S^{-1}
-\end{equation}
-
-%And so on for future state $k+1,...$.
-
-\subsubsection{Nonlinear Observer}
-
-The theory of the state observer was first introduced by Kalman and Bucy for a linear system in a stochastic environment.
-Then Luenberger \cite{david1971introduction} made a general theory of observers for deterministic linear systems, introducing the notions of observer and reduced minimum observer \cite{primbs1996survey}.
-Mahony et al.
-\cite{mahony_nonlinear_2008} have proposed an nonlinear observer called Constant Gain Observer (CGO), termed the explicit complementary filter, that provides attitude estimates as well as gyro bias estimates.
-
-\begin{itemize}
-\item Prediction
-\end{itemize}
-
-First, we propagate the estimate attitude and the gyro bias with equation \ref{prediction_cgo}.
-
-\begin{equation}
-\left\{\begin{array}{l}
-\mathbi{q}_{k}^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro} - \boldsymbol \beta_{k-1}^+)\mathbi{q}_{k-1}^+ \\
-\boldsymbol \beta_k^- = \boldsymbol \beta_{k-1}^+
- \end{array}
-\right.
-\label{prediction_cgo}
-\end{equation}
-
-\begin{itemize}
-\item Correction
-\end{itemize}
-
-The idea is to compute a  constraint  given by a comparison between  an estimate value of a sensor and his real data at a time $k$.
-
-\begin{equation}
-\left\{\begin{array}{l}
-\textbf{v}_{acc}^- = R_(\mathbi{q}_{k}^-)\textbf{a}^a + \boldsymbol  \beta_{acc} \\
-\textbf{v}_{mag}^- = R_(\mathbi{q}_{k}^-)\textbf{m}^a + \boldsymbol  \beta_{mag}
- \end{array}
-\right.
-\end{equation}
-
-All these vectors are normalized.
-
-The '' gain '' is given by,
-
-\begin{equation}
-\boldsymbol \omega_{mes} = -vex(\sum_{i=1}^n \frac{k_i}{2}(\textbf{v}_i(\textbf{v}_i^-)^T -  \textbf{v}_i^-(\textbf{v}_i)^T )) 
-\label{cgo_gain}
-\end{equation}
-
-where the $vex$ operator define by the inverse of the cross product.
-\begin{equation}
-\left\{\begin{array}{l}
-\textbf{a}(\textbf{b})^T -  \textbf{b}(\textbf{a})^T  = [ (\textbf{a} \times \textbf{b}) \times ]\\
-vex([\textbf{a}\times])  = \textbf{a}
- \end{array}
-\right.
-\end{equation}
-
-Then, we the gain is used to correct the predicted attitude $\mathbi{q}_{k}^-$ by the following equation:
-
-\begin{equation}
-\left\{\begin{array}{l}
-\mathbi{q}_{k}^+ = \mathbi{q}_{k}^- + \frac{\Delta T}{2} \Xi(\mathbi{q}_{k}^-) \boldsymbol\omega_{mes}k_p\\
-\boldsymbol \beta_k^+ = \boldsymbol \beta_{k}^-  - \boldsymbol\omega_{mes}k_i
- \end{array}
-\right.
-\end{equation}
-
-One problem of Kalman Filters is to find the initial values in the covariance (matrix $Q$) because it is difficult to relate it to real physical quantities.
-The covariance matrix is not involve in nonlinear observer theory.
-This is one of the reasons there has been an increasing interest for nonlinear observer design the last decades.
-Another reason is their low computational need.
-
-\subsubsection{Methods Based on Solutions of the Wahba's Problem}
-
-The attitude determination problem for vector observation was first formulated as a least squares estimation problem by Wahba \cite{wahba_least_1965} in 1965 :\\
-
-Given the two sets of $n$ vectors $\{ \textbf{r}_{1},...,\textbf{r}_{n} \}$ and $\{ \textbf{b}_{1},...,\textbf{b}_{n} \}$, $n \geqslant 2 $, where each pair $(\textbf{r}_{i},\textbf{b}_{i})$ corresponds to a generalised vector, $\textbf{x}_{i}$, find the proper orthogonal matrix, $A$, which bring the first into the best least squares coincidence with the second.
-That is, find $A$ which minimises the function $J$ defined as:
-
-\begin{equation}
-J(A) = \frac{1}{2} \sum_{i=1}^n a_i||b_i-Ar_i||^2
-\label{Wahba_loss_function}
-\end{equation}
-
-subject to the constraint $A^TA = I$ and $det(A)=1$. The equation (\ref{Wahba_loss_function}) can be written with a quaternion representation which gives:
-
-\begin{equation}
-J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i||b_i-R(\mathbi{q})r_i||^2
-\label{QUEST}
-\end{equation}
-
-The matrix $R(\mathbi{q})$, is the rotation matrix that corresponds to the quaternion $\mathbi{q}$ (equation \ref{quat_to_rot}). The QUEST algorithm are conducted in the following manner. Let
-
-\begin{equation}
-B = \sum_{i=1}^N a_i\textbf{b}_i \textbf{r}_i^T
-\end{equation}
-
-Then, we can rewrite the equation (\ref{QUEST}) as follow
-
-\begin{equation}
-J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i  - tr(R(\mathbi{q})B^T)
-\label{QUEST_bis}
-\end{equation}
-
-Due to the fact that the representation of the attitude matrix is a homogenous quadratic function of $\mathbi{q}$, we can say that
-
-\begin{equation}
-tr(R(\mathbi{q})B^T) = \mathbi{q}^TK\mathbi{q}
-\label{QUEST_ter}
-\end{equation}
-
-where $K$ is the symmetric traceless matrix such as
-
-\begin{equation}
-K = \begin{pmatrix} S-tr(B)I_3 & \textbf{z} \\ \textbf{z}^T & tr(B)
-\end{pmatrix}
-\end{equation}
-
-with
-
-\begin{equation}
-\left\{\begin{array}{l}
-S = B + B^T\\
-\textbf{z} = \sum_ia_i\textbf{b}_i\times\textbf{r}_i
- \end{array}
-\right.
-\end{equation}
-
-The minimisation problem from Equation (\ref{QUEST_bis}) is now a maximisation problem.
-thus, the solution $\mathbi{q}_{opt}$ is the eingen vector correspond to the largest eigen value of $K$ \cite{markley1999estimate}.
-The main problem of QUEST is that it does not take into account of all the previous observations performed since the beginning.
-A recursive QUEST algorithm (REQUEST) proposed by Bar-Itzhack manages this problem.
-According to the REQUEST algorithm, the propagation of $K_k$ is given by
-
-\begin{equation}
-K_k^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro})K_{k-1}^+\Phi_{k|k-1}^T(\boldsymbol\omega_{gyro})
-\end{equation}
-
-We compute $\delta K_k $ as
-
-\begin{equation}
-\delta K_k = \frac{1}{\sum_ia_i}\begin{pmatrix} \delta S_k-\delta\sigma_kI_3 & \delta\textbf{z}_k \\ \delta\textbf{z}_k^T & \delta\sigma_k
-\end{pmatrix}
-\end{equation}
-
-with,
-
-\begin{equation}
-\begin{array}{cc}
-\delta B_k = \sum_i a_i \textbf{b}_i\textbf{r}_i^T & \delta S_k = B_k + B_k^T\\
-\delta \textbf{z}_k = \sum_ia_i\textbf{b}_i\times\textbf{r}_i & \delta \sigma_k = tr(B_k)
-\end{array}
-\end{equation}
-
-Where $\textbf{b}_i$ is the observation vector at time $k$. Then we update the K-matrix.
-
-\begin{equation}
-K_k^+ = \frac{\rho_k m_{k-1}}{m_k}K_k^- + \frac{1}{m_k}\delta K_k
-\end{equation}
-
-A drawback is that the value of $\rho_k$ comes from heuristics.
-Choukroun revisit REQUEST provides an improvement with his Optimal REQUEST \cite{choukroun_novel_2003} where an optimal $\rho_k^*$ is computed which gives better results.
-Here we choose to implement this version of REQUEST.
-Contrary to other methods, the QUEST algorithms do not estimate the gyroscope bias.
-Which can make it less robust in some cases.
-There are also other methods based on Wahba's problem like the Davenport's q-method \cite{weighted1971nasa} proposed in 1971 considered more robust than QUEST.
-Or the method ESOQ proposed by Mortari \cite{mortari1997esoq}  in 1997.
-But in this study we limit our investigation to REQUEST.\\
-
-\subsubsection{Particle Filtering Method}
-
-Particle filtering method is based on Monte-Carlo method\cite{metropolis1949monte}.
-The main idea is to use a large number of samples called particles to estimate a nonlinear function \cite{chen_bayesian_2003} and \cite{terejanu2009tutorial}.
-The set of $N$ particle with associated weight at time $k$ are denoted by $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$.
-Initially, the particle are drawn from a proposal distribution.
-The PF is divided into three step, prediction, update, resampling, that constitute a filter cycle:
-
-\begin{itemize}
-\item Prediction
-\end{itemize}
-
-The first step is to propagate all particles $\{\textbf{x}^{(i)}_{k-1},w^{(i)}_{k-1}\}$ to $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$ with the nonlinear function $f$ while the weight remain unchanged :
-
-\begin{equation}
-\textbf{x}^{(i)}_{k} = f(\textbf{x}^{(i)}_{k-1},\textbf{u}_{k-1}) + \textbf{w}_{k-1}^{(i)}\\
-\end{equation}
-
-where $N$ sample $\textbf{w}_{k-1}^{(i)}$ of the process noise (supposed gaussian in our case) are drawn.
-
-\begin{itemize}
-\item Update/Correction
-\end{itemize}
-
-At the update step, the weights associated with each particle is updated based on the likelihood function $p( \textbf{z}_k| \textbf{x}_k^{(i)})$.
-\begin{equation}
-\left\{ 
-\begin{array}{l}
-w_k^{(i)} = w_{k-1}^{(i)} p( \textbf{z}_k| \textbf{x}_k^{(i)})\\
-\tilde{w}_k^{(i)} = \frac{w_{k}^{(i)}}{\sum_iw_{k}^{(i)}}
-\end{array}
-\right.
-\end{equation}
-
-The weight are normalised such as we have $\sum_i\tilde{w}_k^{(i)}=1$. To finish, the mean is computed with the following equation:
-
-\begin{equation}
- \textbf{x}_k^+ = \sum_{i=1}^N\tilde{w}_k^{(i)} \textbf{x}_k^{(i)}
- \end{equation}
-
-\begin{itemize}
-\item Resampling/Regularisation
-\end{itemize}
-
-A common problem with PF is the degeneracy phenomenon.
-After some iterations, more and more particles will have a weak weight, which make them negligible and useless.
-To overcome this problem, a regularisation step called resampling is processed.
-It is not needed to perform PF but it allows to maintain a good performance relative to degeneracy problem.
-The idea of resampling is to eliminate particles having feeble weight and focus on strong weight particles.
-In order to measure degeneracy, the number of effective samples $N_{eff}$ is computed (equation \ref{N_eff}).
-If  $N_{eff}$ is lower than a threshold, the resampling step is processed.
-
-\begin{equation}
- N_{eff} = \sum_{i=1}^N\frac{1}{(\tilde{w}_k^{(i)})^2}
- \label{N_eff}
- \end{equation}
-
-The choice of the threshold is important.
-Indeed, a high threshold gives $N_{eff} \approx N$, but that does not mean necessarily we have a lot of efficient particles.
-On the contrary, it means that most of the particle are identical making large computations unnecessary.
-Moreover, with a low threshold, many particles are neglected making computations less efficient.
-In general, the threshold is fixed at  $\frac{N}{2}$ or $\frac{N}{3}$.\\
-
-There are several types of resampling, a comparison is made in \cite{douc2005comparison}.
-We took the resampling algorithm described in \cite{arulampalam2002tutorial} because it is the most commun.
-Cheng, Yang and Crassidis have propose an attitute estimation with a particle filter method \cite{cheng_particle_2010}.
-We have also considered the revisited algorithm of Cheng \cite{chang_particle_2014} based on bootstrap filter\cite{gordon1993novel}.
-We have chosen to implement this one because they use the same approach as USQUE \cite{crassidis_unscented_2003} and MEKF\cite{markley2003attitude}.
-We decided to explore the performance of this three algorithms with the same model.
-\\
-
-The particle filter provides an estimate of the nonlinear system without assuming that the noise is Gaussian.
-So it is more robust to any situation, which makes it more performant in the context of a real applications.
-But an important computation time is required and that is a limitation for a real-time application.
-
-\section{Performance Evaluation}
-
-We have access through our designed ground truth to the true attitude $\mathbi{q}^{real}_k$ at time $k$ of the reference frame $\{c\}$ related to $\{a\}$.
-By computation, each method returns an estimate attitude $\mathbi{q}^{est}_k$ at time $k$ (see figure \ref{test_method}).
-
-\begin{equation}
- \delta \mathbi{q}_k = \mathbi{q}^{real}_k \otimes (\mathbi{q}^{est}_k)^{-1} =  \begin{pmatrix}  \delta \textbf{q}_{k} \\ \delta q_{w_k}  \end{pmatrix}
-\end{equation}
-
-The error-MRP vector related to $ \delta \mathbi{q}_k$ is given by
-
-\begin{equation}
- \delta \textbf{p}_k  = 4[\frac{\delta \textbf{q}_{k}  }{(1+\delta q_{w_k})}]
-\end{equation}
-
-Thus, we can calculate the attitude error given by the root mean square (RMS) of $\delta \textbf{p}_k$:
-
-\begin{equation}
-\epsilon_k  = RMS (\delta \textbf{p}_k) = \sqrt{\frac{1}{3}( \delta p_{k_x}^2 +\delta p_{k_y}^2+\delta p_{k_z}^2)  }
-\label{error_definition}
-\end{equation}
-
-The mean of the error is given by :
-
-\begin{equation}
-\bar{\epsilon}  = \frac{1}{n}\sum_{i=1}^n \epsilon_i \\
-\label{mean_var_error_def}
-\end{equation}
-
-With $n$ the number of data at our disposal. We use this information (mean error) for our performance study.
-
-\begin{figure}[!h]
-\centering
-\includegraphics[scale=0.40]{images/test_method.png}
-\caption{Algorithm designed to compute attitude estimation methods' performances}
-\label{test_method}
-\end{figure}
-
-To differentiate the methods, we assigned  some ''scores'' in order to highlight the behaviour of these methods in different configurations. We defined these ''scores'' as:
-
-\begin{align}
-s_1 = \bar{\epsilon}\\
-s_2 = \frac{1}{\bar{\epsilon} \times \bar{\tau}}
-\label{score}
-\end{align}
-
-With $\bar{\epsilon}$ average attitude error (voir \ref{error_definition}) et $\bar{\tau}$ the average computation time. \\
-
-\begin{figure}[!h]
-
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.85\linewidth]{images/histo_s1_add.png}
-  \caption{Mean Attitude Error (score $s_1$)}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=.85\linewidth]{images/histo_s2_add.png}
-  \caption{Ratio Error times Computation Time (score $s_2$)}
-
-\end{subfigure}
-\caption{Scores of methods where noise is AWGN with 3 strength}
-
-\end{figure}
-
-\section{Conclusion}
-
-notre library est trop bien...
-
-\bibliographystyle{ieeetr}
+\input{conclusion.tex}
 

### 2016-05-09 Article/prez_library.tex
adding different file for article

\section{Global Presentation of the Library}

\texttt{OpenMotion} est une librairie open source qui permet a l'utilisateur d'estimer l'orientation du telephone a partir des donnee de l'IMU. La librairie ne supporte pas encore les translations, nous nous sommes concentre sur l'attitude.

\begin{itemize}
\item The North East Down (NED) frame $\{a\}$ system has its origin fixed at the (moving) object center of gravity. The $z$-axis points upward perpendicularly to the tangent plane of the ellipsoid, and the $x$-axis points towards true north (and not the magnetic north). The $y$-axis point towards east.

\vspace{0.1cm}

\item The object-fixed reference frame $\{c\}$ is a moving and rotating coordinate frame that is fixed to the smartphone device. 
\end{itemize}

The library admit admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure~\ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal} .

\begin{figure}[!h]
\centering
\includegraphics[scale=0.55]{images/Schema_situation.png}
\caption{Definition of the scene with 2 frame:  a fixed frame NED (North East Down) noted $\{a\}$ and a moving frame object noted $\{c\}$}
\label{Schema_situation}
\end{figure}

La librairie donne une estimation le quaternion $\mathbi{q}_a^c$

### 2016-05-09 Article/state_of_the_art.tex
adding different file for article

\section{State of the art}

\subsection{Kinematic and quaternion representation}

A quaternion is a four-dimensional vector, defined as

\begin{center}
$ \mathbi{q} = \begin{pmatrix} \textbf{q} \\ q_w \end{pmatrix} $ with $ \textbf{q} =  ( q_x, q_y, q_z)^T $. 
\end{center}

For attitude representation, the quaternion must satisfies a single constraint given by  $|| \mathbi{q} || = 1$.The quaternion kinematics equation in discrete-time is given by :

\begin{equation}
\mathbi{q}_{k+1} = \Phi_{k+1|k}(\boldsymbol\omega)\mathbi{q}_{k}
\label{quat_kine}
\end{equation}

where $\boldsymbol\omega$ is the angular velocity and $\Phi_{k+4|k}$ is the transition matrix from time $k-1$ to $k$ such as:

\begin{equation}
 \Phi_{k+1|k}(\boldsymbol\omega) = \begin{pmatrix}  \cos(0.5||\boldsymbol\omega||\Delta T)I_3 - [\boldsymbol \psi\times] & \boldsymbol\psi \\ -\boldsymbol\psi^T &  \cos(0.5||\boldsymbol\omega||\Delta T)  \end{pmatrix} 
\end{equation}

 with,

\begin{equation}
 \boldsymbol \psi = \sin(0.5||\boldsymbol\omega||\Delta T)\frac{\boldsymbol\omega}{||\boldsymbol\omega||}
\end{equation}

And $[\textbf{v} \times] \in \mathbb{R}^{3\times 3}$ is the skew-symetric matrix of the vector $\textbf{v}$ such as.

\begin{equation}
[\textbf{v} \times] = \begin{pmatrix} 0 & v_z & -v_y \\ -v_z & 0 & v_x \\ v_y & -v_x & 0 \end{pmatrix} 
\label{skewsymmat}
\end{equation}
Then, the attitude matrix $R \in SO(3) $ related to the quaternion $\mathbi{q}$ is given by :

\begin{equation}
R(\mathbi{q}) = (q_w^2 - |\textbf{q}|^2)I + 2 \textbf{q}\textbf{q}^T - 2q_w[\textbf{q} \times]
\label{quat_to_rot}
\end{equation}

Where $SO(3)$ is the special orthogonal group defined by:

\begin{center}
$SO(3) = \{ R \in \mathbb{R}^{ 3 \times 3} | RR^T = R^TR = I, det(R) = 1 \}$
\end{center}

\subsection{Most Common Fusion Methods}

\subsubsection{Kalman Filtering Methods}

Introduced in 1960 by Kalman \cite{kalman_new_1960}, the Kalman filter is an algorithm that uses a series of measurements observed over time, containing noise, in order to estimate unknown variables (state vector). %The goal is to obtain more precise results than those based on a single measurement alone. 
Based on a model it is possible to obtain a more robust estimation of the state vector. Consider the following linear system, described by the difference equation and the observation model with additive noise:

\begin{equation}
\left\{ \begin{array}{l}
\textbf{x}_{k} = F_{k}\textbf{x}_{k-1}+B_{k}\textbf{u}_{k}+G_{k}\textbf{v}_{k}\\
\textbf{z}_{k} = H_{k}\textbf{x}_{k}+D_{k}\textbf{w}_{k}
\end{array} \right.
\label{linear_system}
\end{equation}

with,

\begin{center}
\begin{tabular}{rl}
$\textbf{x}_{k} $ &State vector at time $k$\\
$\textbf{z}_{k} $ &Measurement vector at time $k$\\
$\textbf{u}_{k} $ & Input vector (e.i.: from sensor) at time $k$ \\
$\textbf{v}_{k} $ & State noise vector disturbing the system\\
$\textbf{w}_{k} $ & Observation noise disturbing the measurement\\
\end{tabular}
\end{center}

\vspace{0.2cm}

The figure \ref{State_equation_discrete_case} represents the dynamical system described in \ref{linear_system} . A brief presentation of the algorithm is provided here,  see \cite{terejanu2013discrete} for more details. The algorithm has 3 important steps :\\

\begin{itemize}
\item The first step is the \textbf{prediction} of state $\textbf{x}^-_{k}$ at time $k$ and covariance $P^-_{k}$ is given by :

\begin{equation}
\left\{ \begin{array}{cl}
\textbf{x}^-_{k} = & F_{k}\textbf{x}^+_{k-1} + B_{k}\textbf{u}_{k} \\
P^-_{k} = & F_{k}P^+_{k-1}F^T_{k}+G_{k}Q_{k}G^T_{k}
\end{array}
\right.
\end{equation}

\item The \textbf{innovation} is the difference between predicted measurement and real measurement. It measure up the deviation used by the Kalman gain.
\begin{equation}
\left\{ \begin{array}{cl}
\boldsymbol\nu_{k} = & \textbf{z}_{k}-H_{k}\textbf{x}^-_{k} \\
S_{k} = & R_{k} +H_{k}P^-_{k}H^T_{k}
\end{array}
\right.
\end{equation}

\item And finally, An \textbf{update} is processed to correct the estimate state $\textbf{x}^+_{k} $ and his covariance $P^+_{k}$ with the innovation.

\begin{equation}
\left\{ \begin{array}{cl}
\textbf{x}^+_{k} = & \textbf{x}^-_{k} + K_{k}\boldsymbol\nu_{k} \\
P^+_{k} = &P^-_{k} - K_{k}S_{k}K^T_{k}
\end{array}
\right.
\end{equation}

\end{itemize}

Where $K_k$ is the Kalman gain  representing relative importance of innovation  $\nu_{k}$.

\begin{equation}
K_{k}=P^-_{k}H^T_{k}S^{-1}
\end{equation}

%And so on for future state $k+1,...$.

\subsubsection{Nonlinear Observer}

The theory of the state observer was first introduced by Kalman and Bucy for a linear system in a stochastic environment. Then Luenberger \cite{david1971introduction} made a general theory of observers for deterministic linear systems, introducing the notions of observer and reduced minimum observer \cite{primbs1996survey}. Mahony et al. \cite{mahony_nonlinear_2008} have proposed an nonlinear observer called Constant Gain Observer (CGO), termed the explicit complementary filter, that provides attitude estimates as well as gyro bias estimates.

\begin{itemize}
\item Prediction
\end{itemize}

First, we propagate the estimate attitude and the gyro bias with equation \ref{prediction_cgo}.

\begin{equation}
\left\{\begin{array}{l}
\mathbi{q}_{k}^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro} - \boldsymbol \beta_{k-1}^+)\mathbi{q}_{k-1}^+ \\
\boldsymbol \beta_k^- = \boldsymbol \beta_{k-1}^+
 \end{array}
\right.
\label{prediction_cgo}
\end{equation}

\begin{itemize}
\item Correction
\end{itemize}

The idea is to compute a  constraint  given by a comparison between  an estimate value of a sensor and his real data at a time $k$.

\begin{equation}
\left\{\begin{array}{l}
\textbf{v}_{acc}^- = R_(\mathbi{q}_{k}^-)\textbf{a}^a + \boldsymbol  \beta_{acc} \\
\textbf{v}_{mag}^- = R_(\mathbi{q}_{k}^-)\textbf{m}^a + \boldsymbol  \beta_{mag}
 \end{array}
\right.
\end{equation}

All these vectors are normalized.

The '' gain '' is given by,

\begin{equation}
\boldsymbol \omega_{mes} = -vex(\sum_{i=1}^n \frac{k_i}{2}(\textbf{v}_i(\textbf{v}_i^-)^T -  \textbf{v}_i^-(\textbf{v}_i)^T )) 
\label{cgo_gain}
\end{equation}

where the $vex$ operator define by the inverse of the cross product.
\begin{equation}
\left\{\begin{array}{l}
\textbf{a}(\textbf{b})^T -  \textbf{b}(\textbf{a})^T  = [ (\textbf{a} \times \textbf{b}) \times ]\\
vex([\textbf{a}\times])  = \textbf{a}
 \end{array}
\right.
\end{equation}

Then, we the gain is used to correct the predicted attitude $\mathbi{q}_{k}^-$ by the following equation:

\begin{equation}
\left\{\begin{array}{l}
\mathbi{q}_{k}^+ = \mathbi{q}_{k}^- + \frac{\Delta T}{2} \Xi(\mathbi{q}_{k}^-) \boldsymbol\omega_{mes}k_p\\
\boldsymbol \beta_k^+ = \boldsymbol \beta_{k}^-  - \boldsymbol\omega_{mes}k_i
 \end{array}
\right.
\end{equation}

One problem of Kalman Filters is to find the initial values in the covariance (matrix $Q$) because it is difficult to relate it to real physical quantities. The covariance matrix is not involve in nonlinear observer theory. This is one of the reasons there has been an increasing interest for nonlinear observer design the last decades. Another reason is their low computational need.

\subsubsection{Methods Based on Solutions of the Wahba's Problem}

The attitude determination problem for vector observation was first formulated as a least squares estimation problem by Wahba \cite{wahba_least_1965} in 1965 :\\

Given the two sets of $n$ vectors $\{ \textbf{r}_{1},...,\textbf{r}_{n} \}$ and $\{ \textbf{b}_{1},...,\textbf{b}_{n} \}$, $n \geqslant 2 $, where each pair $(\textbf{r}_{i},\textbf{b}_{i})$ corresponds to a generalised vector, $\textbf{x}_{i}$, find the proper orthogonal matrix, $A$, which bring the first into the best least squares coincidence with the second. That is, find $A$ which minimises the function $J$ defined as:

\begin{equation}
J(A) = \frac{1}{2} \sum_{i=1}^n a_i||b_i-Ar_i||^2
\label{Wahba_loss_function}
\end{equation}

subject to the constraint $A^TA = I$ and $det(A)=1$. The equation (\ref{Wahba_loss_function}) can be written with a quaternion representation which gives:

\begin{equation}
J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i||b_i-R(\mathbi{q})r_i||^2
\label{QUEST}
\end{equation}

The matrix $R(\mathbi{q})$, is the rotation matrix that corresponds to the quaternion $\mathbi{q}$ (equation \ref{quat_to_rot}). The QUEST algorithm are conducted in the following manner. Let

\begin{equation}
B = \sum_{i=1}^N a_i\textbf{b}_i \textbf{r}_i^T
\end{equation}

Then, we can rewrite the equation (\ref{QUEST}) as follow

\begin{equation}
J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i  - tr(R(\mathbi{q})B^T)
\label{QUEST_bis}
\end{equation}

Due to the fact that the representation of the attitude matrix is a homogenous quadratic function of $\mathbi{q}$, we can say that

\begin{equation}
tr(R(\mathbi{q})B^T) = \mathbi{q}^TK\mathbi{q}
\label{QUEST_ter}
\end{equation}

where $K$ is the symmetric traceless matrix such as

\begin{equation}
K = \begin{pmatrix} S-tr(B)I_3 & \textbf{z} \\ \textbf{z}^T & tr(B)
\end{pmatrix}
\end{equation}

with

\begin{equation}
\left\{\begin{array}{l}
S = B + B^T\\
\textbf{z} = \sum_ia_i\textbf{b}_i\times\textbf{r}_i
 \end{array}
\right.
\end{equation}

The minimisation problem from Equation (\ref{QUEST_bis}) is now a maximisation problem. thus, the solution $\mathbi{q}_{opt}$ is the eingen vector correspond to the largest eigen value of $K$ \cite{markley1999estimate}. The main problem of QUEST is that it does not take into account of all the previous observations performed since the beginning. A recursive QUEST algorithm (REQUEST) proposed by Bar-Itzhack manages this problem. According to the REQUEST algorithm, the propagation of $K_k$ is given by

\begin{equation}
K_k^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro})K_{k-1}^+\Phi_{k|k-1}^T(\boldsymbol\omega_{gyro})
\end{equation}

We compute $\delta K_k $ as

\begin{equation}
\delta K_k = \frac{1}{\sum_ia_i}\begin{pmatrix} \delta S_k-\delta\sigma_kI_3 & \delta\textbf{z}_k \\ \delta\textbf{z}_k^T & \delta\sigma_k
\end{pmatrix}
\end{equation}

with,

\begin{equation}
\begin{array}{cc}
\delta B_k = \sum_i a_i \textbf{b}_i\textbf{r}_i^T & \delta S_k = B_k + B_k^T\\
\delta \textbf{z}_k = \sum_ia_i\textbf{b}_i\times\textbf{r}_i & \delta \sigma_k = tr(B_k)
\end{array}
\end{equation}

Where $\textbf{b}_i$ is the observation vector at time $k$. Then we update the K-matrix.

\begin{equation}
K_k^+ = \frac{\rho_k m_{k-1}}{m_k}K_k^- + \frac{1}{m_k}\delta K_k
\end{equation}

A drawback is that the value of $\rho_k$ comes from heuristics. Choukroun revisit REQUEST provides an improvement with his Optimal REQUEST \cite{choukroun_novel_2003} where an optimal $\rho_k^*$ is computed which gives better results. Here we choose to implement this version of REQUEST. Contrary to other methods, the QUEST algorithms do not estimate the gyroscope bias. Which can make it less robust in some cases.
There are also other methods based on Wahba's problem like the Davenport's q-method \cite{weighted1971nasa} proposed in 1971 considered more robust than QUEST. Or the method ESOQ proposed by Mortari \cite{mortari1997esoq}  in 1997. But in this study we limit our investigation to REQUEST.\\

\subsubsection{Particle Filtering Method}

Particle filtering method is based on Monte-Carlo method\cite{metropolis1949monte}. The main idea is to use a large number of samples called particles to estimate a nonlinear function \cite{chen_bayesian_2003} and \cite{terejanu2009tutorial}. The set of $N$ particle with associated weight at time $k$ are denoted by $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$. Initially, the particle are drawn from a proposal distribution. The PF is divided into three step, prediction, update, resampling, that constitute a filter cycle:

\begin{itemize}
\item Prediction
\end{itemize}

The first step is to propagate all particles $\{\textbf{x}^{(i)}_{k-1},w^{(i)}_{k-1}\}$ to $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$ with the nonlinear function $f$ while the weight remain unchanged :

\begin{equation}
\textbf{x}^{(i)}_{k} = f(\textbf{x}^{(i)}_{k-1},\textbf{u}_{k-1}) + \textbf{w}_{k-1}^{(i)}\\
\end{equation}

where $N$ sample $\textbf{w}_{k-1}^{(i)}$ of the process noise (supposed gaussian in our case) are drawn.

\begin{itemize}
\item Update/Correction
\end{itemize}

At the update step, the weights associated with each particle is updated based on the likelihood function $p( \textbf{z}_k| \textbf{x}_k^{(i)})$.
\begin{equation}
\left\{ 
\begin{array}{l}
w_k^{(i)} = w_{k-1}^{(i)} p( \textbf{z}_k| \textbf{x}_k^{(i)})\\
\tilde{w}_k^{(i)} = \frac{w_{k}^{(i)}}{\sum_iw_{k}^{(i)}}
\end{array}
\right.
\end{equation}

The weight are normalised such as we have $\sum_i\tilde{w}_k^{(i)}=1$. To finish, the mean is computed with the following equation:

\begin{equation}
 \textbf{x}_k^+ = \sum_{i=1}^N\tilde{w}_k^{(i)} \textbf{x}_k^{(i)}
 \end{equation}

\begin{itemize}
\item Resampling/Regularisation
\end{itemize}

A common problem with PF is the degeneracy phenomenon. After some iterations, more and more particles will have a weak weight, which make them negligible and useless.  To overcome this problem, a regularisation step called resampling is processed. It is not needed to perform PF but it allows to maintain a good performance relative to degeneracy problem. The idea of resampling is to eliminate particles having feeble weight and focus on strong weight particles. In order to measure degeneracy, the number of effective samples $N_{eff}$ is computed (equation \ref{N_eff}). If  $N_{eff}$ is lower than a threshold, the resampling step is processed.

\begin{equation}
 N_{eff} = \sum_{i=1}^N\frac{1}{(\tilde{w}_k^{(i)})^2}
 \label{N_eff}
 \end{equation}

The choice of the threshold is important. Indeed, a high threshold gives $N_{eff} \approx N$, but that does not mean necessarily we have a lot of efficient particles. On the contrary, it means that most of the particle are identical making large computations unnecessary. Moreover, with a low threshold, many particles are neglected making computations less efficient. In general, the threshold is fixed at  $\frac{N}{2}$ or $\frac{N}{3}$.\\

There are several types of resampling, a comparison is made in \cite{douc2005comparison}. We took the resampling algorithm described in \cite{arulampalam2002tutorial} because it is the most commun. Cheng, Yang and Crassidis have propose an attitute estimation with a particle filter method \cite{cheng_particle_2010}. We have also considered the revisited algorithm of Cheng \cite{chang_particle_2014} based on bootstrap filter\cite{gordon1993novel}. We have chosen to implement this one because they use the same approach as USQUE \cite{crassidis_unscented_2003} and MEKF\cite{markley2003attitude}. We decided to explore the performance of this three algorithms with the same model. \\

The particle filter provides an estimate of the nonlinear system without assuming that the noise is Gaussian. So it is more robust to any situation, which makes it more performant in the context of a real applications. But an important computation time is required and that is a limitation for a real-time application.

\section{Performance Evaluation}

We have access through our designed ground truth to the true attitude $\mathbi{q}^{real}_k$ at time $k$ of the reference frame $\{c\}$ related to $\{a\}$.
By computation, each method returns an estimate attitude $\mathbi{q}^{est}_k$ at time $k$ (see figure \ref{test_method}).

\begin{equation}
 \delta \mathbi{q}_k = \mathbi{q}^{real}_k \otimes (\mathbi{q}^{est}_k)^{-1} =  \begin{pmatrix}  \delta \textbf{q}_{k} \\ \delta q_{w_k}  \end{pmatrix}
\end{equation}

The error-MRP vector related to $ \delta \mathbi{q}_k$ is given by

\begin{equation}
 \delta \textbf{p}_k  = 4[\frac{\delta \textbf{q}_{k}  }{(1+\delta q_{w_k})}]
\end{equation}

Thus, we can calculate the attitude error given by the root mean square (RMS) of $\delta \textbf{p}_k$:

\begin{equation}
\epsilon_k  = RMS (\delta \textbf{p}_k) = \sqrt{\frac{1}{3}( \delta p_{k_x}^2 +\delta p_{k_y}^2+\delta p_{k_z}^2)  }
\label{error_definition}
\end{equation}

The mean of the error is given by :

\begin{equation}
\bar{\epsilon}  = \frac{1}{n}\sum_{i=1}^n \epsilon_i \\
\label{mean_var_error_def}
\end{equation}

With $n$ the number of data at our disposal. We use this information (mean error) for our performance study.

\begin{figure}[!h]
\centering
\includegraphics[scale=0.40]{images/test_method.png}
\caption{Algorithm designed to compute attitude estimation methods' performances}
\label{test_method}
\end{figure}

To differentiate the methods, we assigned  some ''scores'' in order to highlight the behaviour of these methods in different configurations. We defined these ''scores'' as:

\begin{align}
s_1 = \bar{\epsilon}\\
s_2 = \frac{1}{\bar{\epsilon} \times \bar{\tau}}
\label{score}
\end{align}

With $\bar{\epsilon}$ average attitude error (voir \ref{error_definition}) et $\bar{\tau}$ the average computation time. \\
\begin{figure}[!h]

\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[scale=0.28]{images/histo_s1_add.png}
\caption{Scores $s_1$}
\label{histo_s1_add}
\end{subfigure}

\begin{subfigure}{.5\textwidth}
\centering
\includegraphics[scale=0.28]{images/histo_s2_add.png}
\caption{Scores $s_2$ }
\label{histo_s2_add}
\end{subfigure}
\caption{Scores $s_1$ (accuracy) and $s_2$ (accuracy and computation time ratio) of methods where noise is AWGN with 3 strength (weak, moderate, strong)}
\end{figure}

### 2016-05-09 Article/prez_library.tex
creating archi for article smartphone journal

@@
 
-\texttt{OpenMotion} est une librairie open source qui permet a l'utilisateur d'estimer l'orientation du telephone a partir des donnee de l'IMU.
-La librairie ne supporte pas encore les translations, nous nous sommes concentre sur l'attitude.
+Voici les 2 reference frame utiliser pour la creation de la librairie
 
@@
 
-\begin{figure}[!h]
+\begin{figure}
 \centering
-\includegraphics[scale=0.55]{images/Schema_situation.png}
+\includegraphics[scale=0.65]{images/Schema_situation.png}
 \caption{Definition of the scene with 2 frame:  a fixed frame NED (North East Down) noted $\{a\}$ and a moving frame object noted $\{c\}$}
@@
 
-La librairie donne une estimation le quaternion $\mathbi{q}_a^c$
+\texttt{OpenMotion} est une librairie open source qui permet a l'utilisateur d'estimer l'orientation du telephone (representer.
+En utilisant uniquement les donnee de l'IMU integre au telephone.
+La librairie ne supporte pas encore les translations, nous nous sommes concentre sur l'attitude.

### 2016-05-10 Article/main.tex
creating archi for article smartphone journal
modif main.tex

@@
 \usepackage{subcaption}
+\usepackage{mathrsfs}
+\usepackage{listings}
+\usepackage{color}
+
+%\def\includegraphic{}
+%\def\includegraphics{}
 
@@
 \startlocaldefs
+
+%Definition d un environnement remarque auto-indexe
+
+\newcounter{rmq}[section]
+\setcounter{rmq}{0}
+\newenvironment{remarque}{\addtocounter{rmq}{1}\textbf{Remarque \thermq:}}
 
@@
 \makeatother
+
+% colors
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+\definecolor{mauve}{rgb}{0.58,0,0.82}
+
+% code
+\lstset{frame=tb,
+  language=C++,
+  aboveskip=3mm,
+  belowskip=3mm,
+  showstringspaces=false,
+  columns=flexible,
+  basicstyle={\small\ttfamily},
+  numbers=none,
+  numberstyle=\tiny\color{gray},
+  keywordstyle=\color{blue},
+  commentstyle=\color{dkgreen},
+  stringstyle=\color{mauve},
+  breaklines=true,
+  breakatwhitespace=true,
+  tabsize=3
+}
 
@@
 \address[id=aff1]{%                           % unique id
-  \orgname{IPAL (Sorbonne UPMC, CNRS, Astar, NUS, UJF, IMT)}, % university, etc
+  \orgname{Image Pervasive Access Lab (Sorbonne UPMC, CNRS, Astar, NUS, UJF, IMT)}, % university, etc
   \city{Singapore}

### 2016-05-10 Article/prez_library.tex
modif main.tex

@@
 
+\texttt{OpenMotion} est une librairie open source qui.
+En utilisant uniquement les donnee de l'IMU integre au telephone.
+La librairie ne supporte pas encore les translations, nous nous sommes concentre sur l'attitude.
+
 Voici les 2 reference frame utiliser pour la creation de la librairie
+
+\vspace{0.1cm}
 
@@
 
-\item The object-fixed reference frame $\{c\}$ is a moving and rotating coordinate frame that is fixed to the smartphone device. 
+\item The object-fixed reference frame $\{c\}$ corresponding to the IMU device. It is a moving and rotating coordinate frame. 
 \end{itemize}
 
-The library admit admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure~\ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal} .
+\vspace{0.1cm}
+
+The library admit admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure\ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal} .
 
@@
 
-\texttt{OpenMotion} est une librairie open source qui permet a l'utilisateur d'estimer l'orientation du telephone (representer.
-En utilisant uniquement les donnee de l'IMU integre au telephone.
-La librairie ne supporte pas encore les translations, nous nous sommes concentre sur l'attitude.
+La librairie permet aussi l'utilisation de different algo de fusion selon le choix de l'interesse. 
+Ainsi que le type de sortie (quat, rotation matrix, euler angle, rodriguez etc...)
+et elle n'estime que l'orientation par rapport a NED. d'autre transformation sont necessaire pour passer dans d'autre reference frame.
+
+\begin{lstlisting}
+// om_exemple.cpp
+#include <om.h>
+
+using namespace om;
+
+void main(int argc, char** argv){
+
+	// 
+	SensorFusionManager manager;
+
+	//if calibration is necessary
+	manager.calibration();
+
+	// while IMU send data
+	while( isData ){
+
+		// set all IMU data at time t
+		manager.setGyroscopeData(gyro_x,gyro_y,gyro_z);	
+		manager.setAccelerometerData(acc_x,acc_y,acc_z);
+		manager.setMagnetometerData(mag_x,mag_y,mag_z);
+
+		// get different kind of output according to user's choices
+
+		// quaternion
+		Quaternion q_est = manager.getAttitude<Quaternion>();
+
+		//  rotation matrix
+		Matrix R = manager.getAttitude<Matrix>();
+
+		//  axis angle
+		AxisAngle axisAngle = manager.getAttitude<Matrix>(); 
+		...
+	}
+}
+\end{lstlisting}
+
+We are also working on the establishment of a dynamic sensor calibration process on all component of the IMU.
+We insist on the dynamic aspect.
+Indeed, firstly, it allows  the library to be adaptable to any kind of IMU, but also to improve the performance without requesting some effort from the user.

### 2016-05-13 Article/prez_library.tex
modif article journal smartphone

--- 2016-05-09 Article/prez_library.tex
@@
 
-Voici les 2 reference frame utiliser pour la creation de la librairie
+\texttt{OpenMotion} is a open source library mainly aimed for real-time attitude estimation.
+Our approach is based on data fusion technics using only the output of a IMU (composed by a 3D gyroscope,  3D accelerometer and a 3D magnetometer).
+Among several methods proposed by the library, the user has the possibilities to choose which methods fits the best to his expectation according to its performance.
+Here the 2 reference frame to remember.
+
+\vspace{0.1cm}
 
@@
 
-\item The object-fixed reference frame $\{c\}$ is a moving and rotating coordinate frame that is fixed to the smartphone device. 
+\item The object-fixed reference frame $\{c\}$ corresponding to the IMU device. It is a moving and rotating coordinate frame. 
 \end{itemize}
 
-The library admit admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure~\ref{Schema_situation}). It is a cyclopean approximation \cite{ouarti2008multimodal} .
+\vspace{0.1cm}
+
+For simplicity reasons, the library admit that all the components of the IMU belong to the same reference frame $\{c\}$ (ref figure \ref{Schema_situation}).
+It is a cyclopean approximation \cite{ouarti2008multimodal}.
+Moreover, the user has to interface the sends of data from the device to the library because the library does not take care of it.
+\texttt{OpenMotion} provide an estimation of the orientation of $\{c\}$ relative to $\{a\}$.
+The choice of the output type (quaternion, rotation matrix, euler angle) is according to the user needs.
 
@@
 
-\texttt{OpenMotion} est une librairie open source qui permet a l'utilisateur d'estimer l'orientation du telephone (representer.
-En utilisant uniquement les donnee de l'IMU integre au telephone.
-La librairie ne supporte pas encore les translations, nous nous sommes concentre sur l'attitude.
+%
+%\begin{lstlisting}[caption=Exemple of utilisation in C++,label=code]
+%// om_exemple.cpp
+%#include <om.h>
+%
+%using namespace om;
+%
+%void main(int argc, char** argv){
+%
+%	// 
+%	SensorFusionManager manager;
+%
+%	//if calibration is necessary
+%	manager.calibration();
+%	
+%	// while IMU send data
+%	while( isData ){
+%
+%		// set all IMU data at time t
+%		manager.setGyroscopeData(gyro_x,gyro_y,gyro_z);	
+%		manager.setAccelerometerData(acc_x,acc_y,acc_z);
+%		manager.setMagnetometerData(mag_x,mag_y,mag_z);
+%		
+%		// get different kind of output according to user's choices
+%		
+%		// quaternion
+%		Quaternion q_est = manager.getAttitude<Quaternion>(); 
+%		
+%		//  rotation matrix
+%		Matrix R = manager.getAttitude<Matrix>(); 
+%		
+%		//  axis angle
+%		AxisAngle axisAngle = manager.getAttitude<Matrix>(); 
+%		...
+%	}
+%}
+%
+%\end{lstlisting}
+
+We are also working on the establishment of a dynamic sensor calibration process on all component of the IMU.
+We insist on the dynamic aspect.
+Indeed, firstly, it allows  the library to be adaptable to any kind of IMU, but also to improve the performance without requesting some effort from the user.

### 2016-05-13 Article/state_of_the_art.tex
modif article journal smartphone

@@
 
-\subsection{Kinematic and quaternion representation}
+\subsection{Quaternion representation}
 
@@
 
-For attitude representation, the quaternion must satisfies a single constraint given by  $|| \mathbi{q} || = 1$.The quaternion kinematics equation in discrete-time is given by :
+For attitude representation, the quaternion must satisfies a single constraint given by  $|| \mathbi{q}  || = 1$.
 
+Then, the attitude matrix $R \in SO(3) $ related to the quaternion $\mathbi{q}$ is given by :
+\vspace{-0.15cm}
 \begin{equation}
-\mathbi{q}_{k+1} = \Phi_{k+1|k}(\boldsymbol\omega)\mathbi{q}_{k}
-\label{quat_kine}
+R(\mathbi{q}) = (q_w^2 - |\textbf{q}|^2)I + 2 \textbf{q}\textbf{q}^T - 2q_w[\textbf{q} \times]
+\label{quat_to_rot}
 \end{equation}
 
-where $\boldsymbol\omega$ is the angular velocity and $\Phi_{k+4|k}$ is the transition matrix from time $k-1$ to $k$ such as:
-
-\begin{equation}
- \Phi_{k+1|k}(\boldsymbol\omega) = \begin{pmatrix}  \cos(0.5||\boldsymbol\omega||\Delta T)I_3 - [\boldsymbol \psi\times] & \boldsymbol\psi \\ -\boldsymbol\psi^T &  \cos(0.5||\boldsymbol\omega||\Delta T)  \end{pmatrix}
-\end{equation}
-
- with,
-
-\begin{equation}
- \boldsymbol \psi = \sin(0.5||\boldsymbol\omega||\Delta T)\frac{\boldsymbol\omega}{||\boldsymbol\omega||}
-\end{equation}
+Where $SO(3)$ is the special orthogonal group defined by:
+\vspace{-0.15cm}
+$$SO(3) = \{ R \in \mathbb{R}^{ 3 \times 3} | RR^T = R^TR = I, det(R) = 1 \}$$
 
@@
 \end{equation}
-Then, the attitude matrix $R \in SO(3) $ related to the quaternion $\mathbi{q}$ is given by :
-
-\begin{equation}
-R(\mathbi{q}) = (q_w^2 - |\textbf{q}|^2)I + 2 \textbf{q}\textbf{q}^T - 2q_w[\textbf{q} \times]
-\label{quat_to_rot}
-\end{equation}
-
-Where $SO(3)$ is the special orthogonal group defined by:
-
-\begin{center}
-$SO(3) = \{ R \in \mathbb{R}^{ 3 \times 3} | RR^T = R^TR = I, det(R) = 1 \}$
-\end{center}
 
@@
 
-\subsubsection{Kalman Filtering Methods}
+\subsubsection{Kalman Filter}
 
 Introduced in 1960 by Kalman \cite{kalman_new_1960}, the Kalman filter is an algorithm that uses a series of measurements observed over time, containing noise, in order to estimate unknown variables (state vector).
-%The goal is to obtain more precise results than those based on a single measurement alone.
 Based on a model it is possible to obtain a more robust estimation of the state vector.
@@
 \left\{ \begin{array}{l}
-\textbf{x}_{k} = F_{k}\textbf{x}_{k-1}+B_{k}\textbf{u}_{k}+G_{k}\textbf{v}_{k}\\
+\textbf{x}_{k+1} = F_{k}\textbf{x}_{k}+B_{k}\textbf{u}_{k}+G_{k}\textbf{v}_{k}\\
 \textbf{z}_{k} = H_{k}\textbf{x}_{k}+D_{k}\textbf{w}_{k}
@@
 
-The figure \ref{State_equation_discrete_case} represents the dynamical system described in \ref{linear_system} .
-A brief presentation of the algorithm is provided here,  see \cite{terejanu2013discrete} for more details.
-The algorithm has 3 important steps :\\
+A brief presentation of the algorithm is provided here,  see \cite{terejanu2013discrete} for more details. The algorithm has 3 important steps :\\
 
 \begin{itemize}
-\item The first step is the \textbf{prediction} of state $\textbf{x}^-_{k}$ at time $k$ and covariance $P^-_{k}$ is given by :
+\item The first step is the \textbf{prediction} of state $\textbf{x}^-_{k+1}$ at time $k+1$ and covariance $P^-_{k+1}$ is given by :
 
@@
 \left\{ \begin{array}{cl}
-\textbf{x}^-_{k} = & F_{k}\textbf{x}^+_{k-1} + B_{k}\textbf{u}_{k} \\
-P^-_{k} = & F_{k}P^+_{k-1}F^T_{k}+G_{k}Q_{k}G^T_{k}
+\textbf{x}^-_{k+1} = & F_{k}\textbf{x}^+_{k} + B_{k}\textbf{u}_{k} \\
+P^-_{k+1} = & F_{k}P^+_{k}F^T_{k}+G_{k}Q_{k}G^T_{k}
 \end{array}
@@
 \left\{ \begin{array}{cl}
-\boldsymbol\nu_{k} = & \textbf{z}_{k}-H_{k}\textbf{x}^-_{k} \\
-S_{k} = & R_{k} +H_{k}P^-_{k}H^T_{k}
+\boldsymbol\nu_{k+1} = & \textbf{z}_{k+1}-H_{k+1}\textbf{x}^-_{k+1} \\
+\mathcal{S} = & R_{k+1} +H_{k+1}P^-_{k+1}H^T_{k+1}
 \end{array}
@@
 
-\item And finally, An \textbf{update} is processed to correct the estimate state $\textbf{x}^+_{k} $ and his covariance $P^+_{k}$ with the innovation.
+\item And finally, An \textbf{update} is processed to correct the estimate state $\textbf{x}^+_{k+1} $ and his covariance $P^+_{k+1}$ with the innovation.
 
@@
 \left\{ \begin{array}{cl}
-\textbf{x}^+_{k} = & \textbf{x}^-_{k} + K_{k}\boldsymbol\nu_{k} \\
-P^+_{k} = &P^-_{k} - K_{k}S_{k}K^T_{k}
+\textbf{x}^+_{k+1} = & \textbf{x}^-_{k+1} + \mathcal{K}_{k+1}\boldsymbol\nu_{k+1} \\
+P^+_{k+1} = &P^-_{k+1} - \mathcal{K}_{k+1}S_{k+1}\mathcal{K}^T_{k+1}
 \end{array}
@@
 
-Where $K_k$ is the Kalman gain  representing relative importance of innovation  $\nu_{k}$.
+Where $\mathcal{K}_{k+1}$ is the Kalman gain  representing relative importance of innovation  $\nu_{k+1}$.
 
 \begin{equation}
-K_{k}=P^-_{k}H^T_{k}S^{-1}
+\mathcal{K}_{k+1}=P^-_{k+1}H^T_{k+1}\mathcal{S}^{-1}
 \end{equation}
 
-%And so on for future state $k+1,...$.
+The Kalman filter provides an optimal estimation for linear systems.
+Nevertheless, most of studied dynamical systems are nonlinear (attitude estimation, pendula, ...
+).
+Thus, extended Kalman Filter approach  was proposed for nonlinear systems \cite{larson1967application, larson1967precomputation,terejanu2008extended}.
+
+Here, we have chosen to use quaternions for the attitude representation because of its singular-free specificity.
+But special care must be taken when designing a quaternion based extended Kalman filter.
+Because of the constraint on the quaternions, the covariance matrix $P$ can be singular \cite{vik2009integrated} which is problematic for the numerical stability.
+There exist several methods to deal with this issue.
+Multiplicative Extended Kalman Filter (MEKF)\cite{markley2003attitude} and the Additive Extended Kalman Filter (AEKF) \cite{bar1991quaternion} are designed to tackle this problem.The difference between AEKF and MEKF is discussed in \cite{markley_multiplicative_2004}.
+For the library, we choose to implement MEKF.
+
+On the other hand, the Unscented Kalman Filter method was proposed by Julier and Uhlmann \cite{julier_new_1995} in 1995 .
+It belongs to a class of filters called Sigma-Point Kalman Filters.
+Instead of using the Taylor series expension, UKF use an unscented transformation \cite{uhlmann1995dynamic} to approximate the nonlinear projection of mean and covariance of a probability distribution.
+For the library, we choose to implement the filter proposed by Crassidis et al.
+called UnScented QUaternion Estimator (USQUE) \cite{crassidis_unscented_2003}, based on UKF.
+They used the same representation as MEKF.
 
@@
 \cite{mahony_nonlinear_2008} have proposed an nonlinear observer called Constant Gain Observer (CGO), termed the explicit complementary filter, that provides attitude estimates as well as gyro bias estimates.
-
-\begin{itemize}
-\item Prediction
-\end{itemize}
-
-First, we propagate the estimate attitude and the gyro bias with equation \ref{prediction_cgo}.
-
-\begin{equation}
-\left\{\begin{array}{l}
-\mathbi{q}_{k}^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro} - \boldsymbol \beta_{k-1}^+)\mathbi{q}_{k-1}^+ \\
-\boldsymbol \beta_k^- = \boldsymbol \beta_{k-1}^+
- \end{array}
-\right.
-\label{prediction_cgo}
-\end{equation}
-
-\begin{itemize}
-\item Correction
-\end{itemize}
-
-The idea is to compute a  constraint  given by a comparison between  an estimate value of a sensor and his real data at a time $k$.
-
-\begin{equation}
-\left\{\begin{array}{l}
-\textbf{v}_{acc}^- = R_(\mathbi{q}_{k}^-)\textbf{a}^a + \boldsymbol  \beta_{acc} \\
-\textbf{v}_{mag}^- = R_(\mathbi{q}_{k}^-)\textbf{m}^a + \boldsymbol  \beta_{mag}
- \end{array}
-\right.
-\end{equation}
-
-All these vectors are normalized.
-
-The '' gain '' is given by,
-
-\begin{equation}
-\boldsymbol \omega_{mes} = -vex(\sum_{i=1}^n \frac{k_i}{2}(\textbf{v}_i(\textbf{v}_i^-)^T -  \textbf{v}_i^-(\textbf{v}_i)^T )) 
-\label{cgo_gain}
-\end{equation}
-
-where the $vex$ operator define by the inverse of the cross product.
-\begin{equation}
-\left\{\begin{array}{l}
-\textbf{a}(\textbf{b})^T -  \textbf{b}(\textbf{a})^T  = [ (\textbf{a} \times \textbf{b}) \times ]\\
-vex([\textbf{a}\times])  = \textbf{a}
- \end{array}
-\right.
-\end{equation}
-
-Then, we the gain is used to correct the predicted attitude $\mathbi{q}_{k}^-$ by the following equation:
-
-\begin{equation}
-\left\{\begin{array}{l}
-\mathbi{q}_{k}^+ = \mathbi{q}_{k}^- + \frac{\Delta T}{2} \Xi(\mathbi{q}_{k}^-) \boldsymbol\omega_{mes}k_p\\
-\boldsymbol \beta_k^+ = \boldsymbol \beta_{k}^-  - \boldsymbol\omega_{mes}k_i
- \end{array}
-\right.
-\end{equation}
 
@@
 Given the two sets of $n$ vectors $\{ \textbf{r}_{1},...,\textbf{r}_{n} \}$ and $\{ \textbf{b}_{1},...,\textbf{b}_{n} \}$, $n \geqslant 2 $, where each pair $(\textbf{r}_{i},\textbf{b}_{i})$ corresponds to a generalised vector, $\textbf{x}_{i}$, find the proper orthogonal matrix, $A$, which bring the first into the best least squares coincidence with the second.
-That is, find $A$ which minimises the function $J$ defined as:
+That is, find $A \in SO(3)$ which minimises the function $J$ defined as:
 
@@
 
-subject to the constraint $A^TA = I$ and $det(A)=1$. The equation (\ref{Wahba_loss_function}) can be written with a quaternion representation which gives:
+The equation (\ref{Wahba_loss_function}) can be written with a quaternion representation which gives:
 
@@
 A recursive QUEST algorithm (REQUEST) proposed by Bar-Itzhack manages this problem.
-According to the REQUEST algorithm, the propagation of $K_k$ is given by
-
-\begin{equation}
-K_k^- = \Phi_{k|k-1}(\boldsymbol\omega_{gyro})K_{k-1}^+\Phi_{k|k-1}^T(\boldsymbol\omega_{gyro})
-\end{equation}
-
-We compute $\delta K_k $ as
-
-\begin{equation}
-\delta K_k = \frac{1}{\sum_ia_i}\begin{pmatrix} \delta S_k-\delta\sigma_kI_3 & \delta\textbf{z}_k \\ \delta\textbf{z}_k^T & \delta\sigma_k
-\end{pmatrix}
-\end{equation}
-
-with,
-
-\begin{equation}
-\begin{array}{cc}
-\delta B_k = \sum_i a_i \textbf{b}_i\textbf{r}_i^T & \delta S_k = B_k + B_k^T\\
-\delta \textbf{z}_k = \sum_ia_i\textbf{b}_i\times\textbf{r}_i & \delta \sigma_k = tr(B_k)
-\end{array}
-\end{equation}
-
-Where $\textbf{b}_i$ is the observation vector at time $k$. Then we update the K-matrix.
-
-\begin{equation}
-K_k^+ = \frac{\rho_k m_{k-1}}{m_k}K_k^- + \frac{1}{m_k}\delta K_k
-\end{equation}
-
-A drawback is that the value of $\rho_k$ comes from heuristics.
-Choukroun revisit REQUEST provides an improvement with his Optimal REQUEST \cite{choukroun_novel_2003} where an optimal $\rho_k^*$ is computed which gives better results.
-Here we choose to implement this version of REQUEST.
 Contrary to other methods, the QUEST algorithms do not estimate the gyroscope bias.
 Which can make it less robust in some cases.
+
 There are also other methods based on Wahba's problem like the Davenport's q-method \cite{weighted1971nasa} proposed in 1971 considered more robust than QUEST.
 Or the method ESOQ proposed by Mortari \cite{mortari1997esoq}  in 1997.
-But in this study we limit our investigation to REQUEST.\\
+However, for this library, we limit our investigation to REQUEST.\\
 
@@
 
-The first step is to propagate all particles $\{\textbf{x}^{(i)}_{k-1},w^{(i)}_{k-1}\}$ to $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$ with the nonlinear function $f$ while the weight remain unchanged :
+The first step is to propagate all particles $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$ to $\{\textbf{x}^{(i)}_{k+1},w^{(i)}_{k+1}\}$ with the nonlinear function $f$ while the weight remain unchanged :
 
 \begin{equation}
-\textbf{x}^{(i)}_{k} = f(\textbf{x}^{(i)}_{k-1},\textbf{u}_{k-1}) + \textbf{w}_{k-1}^{(i)}\\
+\textbf{x}^{(i)}_{k+1} = f(\textbf{x}^{(i)}_{k},\textbf{u}_{k}) + \textbf{w}_{k}^{(i)}\\
 \end{equation}
 
-where $N$ sample $\textbf{w}_{k-1}^{(i)}$ of the process noise (supposed gaussian in our case) are drawn.
+where $N$ sample $\textbf{w}_{k}^{(i)}$ of the process noise (supposed gaussian in our case) are drawn.
 
@@
 
-At the update step, the weights associated with each particle is updated based on the likelihood function $p( \textbf{z}_k| \textbf{x}_k^{(i)})$.
+At the update step, the weights associated with each particle is updated based on the likelihood function $p( \textbf{z}_{k+1}| \textbf{x}_{k+1}^{(i)})$.
 \begin{equation}
@@
 \begin{array}{l}
-w_k^{(i)} = w_{k-1}^{(i)} p( \textbf{z}_k| \textbf{x}_k^{(i)})\\
-\tilde{w}_k^{(i)} = \frac{w_{k}^{(i)}}{\sum_iw_{k}^{(i)}}
+w_{k+1}^{(i)} = w_{k}^{(i)} p( \textbf{z}_{k+1}| \textbf{x}_{k+1}^{(i)})\\
+\tilde{w}_{k+1}^{(i)} = \frac{w_{k+1}^{(i)}}{\sum_iw_{k+1}^{(i)}}
 \end{array}
@@
 
-The weight are normalised such as we have $\sum_i\tilde{w}_k^{(i)}=1$. To finish, the mean is computed with the following equation:
+The weight are normalised such as we have $\sum_i\tilde{w}_{k+1}^{(i)}=1$. To finish, the mean is computed with the following equation:
 
 \begin{equation}
- \textbf{x}_k^+ = \sum_{i=1}^N\tilde{w}_k^{(i)} \textbf{x}_k^{(i)}
+ \textbf{x}_{k+1}^+ = \sum_{i=1}^N\tilde{w}_{k+1}^{(i)} \textbf{x}_{k+1}^{(i)}
  \end{equation}
@@
 Moreover, with a low threshold, many particles are neglected making computations less efficient.
-In general, the threshold is fixed at  $\frac{N}{2}$ or $\frac{N}{3}$.\\
-
+In general, the threshold is fixed at  $\frac{N}{2}$ or $\frac{N}{3}$.
 There are several types of resampling, a comparison is made in \cite{douc2005comparison}.
@@
 We decided to explore the performance of this three algorithms with the same model.
-\\
-
 The particle filter provides an estimate of the nonlinear system without assuming that the noise is Gaussian.
@@
 
-We have access through our designed ground truth to the true attitude $\mathbi{q}^{real}_k$ at time $k$ of the reference frame $\{c\}$ related to $\{a\}$.
+\subsection{Score definition}
+We have access through our designed ground truth (simulated or from computer vision technics) to the true attitude $\mathbi{q}^{real}_k$ at time $k$ of the reference frame $\{c\}$ related to $\{a\}$.
 By computation, each method returns an estimate attitude $\mathbi{q}^{est}_k$ at time $k$ (see figure \ref{test_method}).
@@
 
+\subsection{Simulated environment}
+
+Our framework has been entierely coded in C++ and all tests have been launched on a Dell Optiplex 9020 with Intel $ Core^{TM}$ i7-4790 CPU at 3.60GHz $\times$ 8.
+All curves were generated with \texttt{R} and \texttt{Matlab}.
+No paralelization or code optimisation have been done, our program turns entirely in sequential manner.
+Our goal is to see the difference of computation duration of all methods without modifying the performance with a specific optimisation.
+
+\begin{table}
+\begin{tabular}{|c|c|c|}
+\hline
+Sensor & Bias instability  & Measurement noise\rule[-2pt]{0pt}{10pt} \\
+\hline
+\hline
+ Gyroscope & $0.316\times 10^{-2} $  $\mu$rad.s$^{-\frac{1}{2}}$ & $ 3.16$ $ \times 10^{-4} $ $\mu$rad.s$^{-\frac{3}{2}}$ \rule[-1.5pt]{0pt}{13pt}\\
+ \hline
+Magnetometer & $0.0023$ mGauss$/\sqrt{Hz}$ & 0.05 mGauss  \rule[-1.5pt]{0pt}{13pt}\\
+ \hline
+Accelerometer &  $0.002$ m.s$^{-2}/\sqrt{Hz} $ & 0.02 m.s$^{-2}$ \rule[-1.5pt]{0pt}{13pt}\\
+ \hline
+\end{tabular}
+\caption{Calibrated IMU data performance specification, Gyrometer, Accelerometer, Magnetometer}
+\label{spec_imu}
+\end{table}
+
+For our Sensor Model, we've used  specification described in table \ref{spec_imu}.
+
 \begin{subfigure}{.5\textwidth}
@@
 \end{figure}
+
+\subsection{Experimental validation}
+
+We have designed a system that can be easily repeatable in laboratory.
+We used a smartphone as our IMU and a calibrated webcam (see figure \ref{situation_validation} ).
+With this system, we have defined our ground-truth as follows.
+A marker is stuck on the smartphone and we calculate an estimate of the attitude of the marker (which represents the smartphone) relative to the reference frame of the webcam $\{w\}$.
+We used the library \texttt{ArUco}  as optical feedback.
+At the same time, we get the data from the IMU of the smartphone to proceed the attitude estimation with our different algorithms.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.35]{images/situation_validation.png}
+\caption{Representation of different reference frame used in this experiment. Here, only $ \text{\mathbi{q}}_w^a $  does not change.}
+\label{situation_validation}
+\end{figure}
+
+At time $t = 0$, the camera and the smartphone has been set such as attitude quaternions $\mathbi{q}^a_w$, $\mathbi{q}^c_a$ et $\mathbi{q}^w_c$ are approximatively known.
+The camera being fixed, $\mathbi{q}^a_w $ does not vary.
+Moreover, from figure \ref{situation_validation}  we have:
+
+\begin{equation}
+\mathbi{q}^c_a =( \mathbi{q}^c_w \otimes \mathbi{q}^a_w )^{-1}
+\end{equation}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.35]{images/output_aruco.png}
+\caption{exemple of output of \texttt{ArUco} with 6 markers. Here the object-fixed reference frame $\{c\}$ is drawn}
+\label{situation_validation}
+\end{figure}
+
+All methods give an estimate of $\mathbi{q}^c_a$ , that are compared with the data from the camera.
+The optical feedback provides a ground truth estimate for this experiment.
+To estimate the error that can occur with the optical feedback.
+We were careful in the experimentation to avoid the occlusion of the marker that is required to estimate the orientation.
+During this experiment, we applied a rotation of 90 degree around the $z_c$  on the smartphone in clockwise direction.
+Then, another rotation of 90 degree around the $z_c$ on the smartphone but in anticlockwise direction to return to the initial state.
+In order to study possible divergence of all methods.

### 2016-05-13 Article/calibration.tex
Calibration.tex + ref biblio
1 page sur la calibration des capteurs + ref associés dans biblio
globale

@@
 
-A toi de jouer Vivien...
+\subsection{Probl\'ematique de la calibration}
+
+	Les syst\`emes embarqu\'e comme les smartphones disposent pour la plupart de capteurs servant au syst\`eme ou \`a l'utilisateur pour se renseigner sur des grandeurs utiles quant \`a l'appareil lui-m\^eme ou bien \`a son environnement.
+Ces capteurs sont donc vari\'es mais on peut notamment distinguer les capteurs constituant une centrale inertielle ou IMU pour Inertial Measurement Unit que sont les acc\'el\'erom\`etres et les gyroscopes 3 axes.
+Ces capteurs sont de plus en plus associ\'es avec un magn\'etom\`etre.
+Ainsi, ces capteurs peuvent constituer un AHRS pour Attitude and Heading Reference System ayant pour but de rep\'erer l'appareil dans l'espace 3D.
+L'acc\'el\'erom\`etre sert ici \`a mesurer l'acc\'el\'eration lin\'eaire, le gyroscope la vitesse angulaire et le magn\'etom\`etre le champ magn\'etique terrestre local et sert dans la plupart des cas de compas.
+Dans le cas d'un smartphone, ces capteurs sont low-cost et de type MEMS pour Micro-Electro-Mechanical Systems.
+D\`es lors, le signal utile \'etant faible, l'erreur induite peut-\^etre du m\^eme ordre de grandeur et donc conduire \`a des impr\'ecisions importantes, notamment dans le cas des AHRS qui font l'objet d'une fusion de donn\'ees.
+Il convient donc d'\'eliminer ou tout du moins de r\'eduire ces erreurs via une calibration initiale de l'ensemble des capteurs utilis\'es.
+
+\subsection{Types et mod\`ele d'erreur}
+
+	Tous les types de capteurs pr\'esentent des biais, facteurs d'\'echelle (scale factors), des erreurs d'inter-couplage (misalignements et cross axis sensitivities) et dans une certaine mesure du bruit al\'eatoire.
+Des erreurs d'ordre sup\'erieur, des erreurs d'inter-couplage d'acc\'el\'eration angulaire ou encore des erreurs sp\'ecifiques comme les softs et irons distorsions pour le magn\'etom\`etre peuvent \'egalement se produire.
+\\
+\\
+Pour l'ensemble des capteurs, un mod\`ele rendant compte des erreurs lors des diff\'erentes mesures peut-\^etre propos\'e :
+
+\begin{displaymath}
+
+	X_i = \textbf{M} \textbf{S} \cdot \left( X_{i,raw} - b \right) + n_i
+
+\end{displaymath}
+
+o\`u $X_i$ est la mesure du capteur comprenant les erreurs, $X_{i, raw}$ la mesure r\'eelle du capteur,  $b$ le biais,  $\textbf{S}$ la matrice rendant compte des scales factor errors, $\textbf{M}$ la matrice des misalignement entre les axes du capteur et $n_i$ le bruit al\'eatoire avec $i \in \{x,y,z\}$.
+\\
+\\
+Les diff\'erents matrices sont alors
+
+\begin{displaymath}
+
+	\textbf{S} =
+
+	\begin{pmatrix}
+
+		s_{xx} & s_{xy} & s_{xz} \\
+
+		s_{yx} & s_{yy} & s_{yz} \\
+
+		s_{zx} & s_{zy} & s_{zz}
+
+	\end{pmatrix}
+
+	\text{ et } \textbf{M} =
+
+	\begin{pmatrix}
+
+	   1 & -\alpha_{yz} & \alpha_{zy} \\
+
+	   0 & 1 & -\alpha_{zx} \\
+
+	   0 & 0 & 1
+
+	\end{pmatrix}
+
+\end{displaymath}
+
+o\`u $\alpha_{ij}$ avec $i,j \in \{x,y,z\}$ est l'angle de misalignement entre l'axe \emph{i} du rep\`ere du capteur avec celui de la plateforme selon l'axe \emph{j} de la plateforme.
+\\
+\\
+Enfin le biais est repr\'esent\'e par un vecteur 
+\begin{displaymath}
+
+	b =
+
+	\begin{pmatrix}
+
+		b_x\\
+
+		b_y \\
+
+		b_z
+
+	\end{pmatrix}
+
+\end{displaymath}
+
+\subsection{M\'ethodes adopt\'ees}
+
+	Une premi\`ere calibration d'usine est effectu\'ee mais s'av\`ere souvent impr\'ecise ou obsol`ete du fait de l'usure des capteurs. D\`es lors, une nouvelle calibration s'av\`ere n\'ecessaire.
+\\
+\\
+	La calibration des capteurs se fera dans un premier temps ind\'ependamment les unes des autres et de mani\`ere statique \emph{i.
+e.} \`a l'initialisation.
+Il est \a noter que nous ne prendrons pas en compte ici l'influence de la temp\erature puisqu'il est d\'ej\`a trait\'e dans la calibration d'usine.
+
+\subsubsection{Acc\'el\'erom\`etre}
+
+	Freescale \copyright ([Freescale]) propose une m\'ethode de calibration lin\'eaire de l'acc\'el\'erom\`etre 3 axes.
+\\
+\\
+	Elle repose sur le fait que la somme des amplitudes des composantes normalis\'ees du champ gravitationnel terrestre sur les 3 axes du capteur doit \^etre \'egale \`a l'amplitude totale de celui-ci quand le capteur est dans un \'etat quasi statique.
+Elle consiste \`a placer l'appareil dans 8 positions donn\'ees pour chaque axe puis d'utiliser une m\'ethode d'optimisation de moindres carr\'es pour d\'eterminer les 12 param\`etres de calibration que sont les scale factors et le biais.
+\\
+\\
+Cette m\'ethode ne permet toutefois pas de d\'eterminer les misalignements.
+
+\subsubsection{Gyroscope}
+
+	\ref{Gyroscope} propose une m\'ethode de calibration du gyroscope.
+Elle repose sur le fait que la somme des mesures selon les 3 axes du gyroscope doit \^etre \'egale \`a la vitesse angulaire en entr\'ee.
+\\
+\\
+	2 \'etapes sont alors n\'ecessaires \`a la calibration du gyroscope.
+La premi\`ere consiste en des rotations horaires et anti-horaires permettant d'\'eliminer la vitesse angulaire de rotation de la terre et le biais du gyroscope en prenant le carr\'e de la norme de la diff\'erence des int\'egrales des mesures dans les 2 sens puis en utilisant une m\'ethode de moindre carr\'e.
+Les scales factors et les angles de misalignements sont ainsi d\'etermin\'es.
+Ensuite, les angles de misalignements et les scales factors sont utilis\'es pour estimer le biais via la diff\'erence des carr\'es de l'\'equation de la somme des vitesses angulaires pour 2 positions diff\'erentes.
+
+\subsubsection{Magn\'etom\`etre}
+
+	Vasconcelos et al.
+\ref{MLE} proposent un algorithme dit de maximum likelihood estimation (MLE) qui utilise l'estimation issue de "Two-Step" (\ref{Two-Step}), \emph{i.
+e.} une m\'ethode d'estimation de moindre carr\'e suivie d'un filtrage \'etendu de Kalman (EKF), comme approximation initiale et qui traite le bruit de mesure comme une distribution normale.
+\\
+\\
+	Le probl\`eme de minimisation  la fonction  de vraisemblance (likelihood) qui est quadratique revient alors \`a d\'eterminer les points de l'ellipso\"{i}de  qui correspondent au mieux aux mesures du capteur.
+\\
+\\
+	Ce probl\`eme est r\'esolu via la m\'ethode du gradient et une m\'ethode de Newton descendante pour des espaces euclidiens \ref{Newton} et la r\`egle de Armijo pour la d\'etermination de la taille du pas.
+Le gradient et la matrice hessienne de la fonction de log-likelihood sont eux calcul\'es analytiquement.
+
+\subsubsection{Calibration on-board}
+
+	Une calibration on-board peut \'egalement s'av\'erer n\'ecessaire du fait d'un besoin de pr\'ecision accrue ou de l'obsolescence de la calibration d'usine due \`a l'usure des capteurs.

### 2016-05-13 Article/calibration.tex
modif calib and sota tex file

--- 2016-05-09 Article/calibration.tex
@@
 
-A toi de jouer Vivien...
+\subsection{Problematique de la calibration}
+
+	Les systemes embarque comme les smartphones disposent pour la plupart de capteurs servant au systeme ou a l'utilisateur pour se renseigner sur des grandeurs utiles quant a l'appareil lui-meme ou bien a son environnement.
+Ces capteurs sont donc varies mais on peut notamment distinguer les capteurs constituant une centrale inertielle ou IMU pour Inertial Measurement Unit que sont les accelerometres et les gyroscopes 3 axes.
+Ces capteurs sont de plus en plus associes avec un magnetometre.
+Ainsi, ces capteurs peuvent constituer un AHRS pour Attitude and Heading Reference System ayant pour but de reperer l'appareil dans l'espace 3D.
+L'accelerometre sert ici a mesurer l'acceleration lineaire, le gyroscope la vitesse angulaire et le magnetometre le champ magnetique terrestre local et sert dans la plupart des cas de compas.
+Dans le cas d'un smartphone, ces capteurs sont low-cost et de type MEMS pour Micro-Electro-Mechanical Systems.
+Des lors, le signal utile etant faible, l'erreur induite peut-etre du meme ordre de grandeur et donc conduire a des imprecisions importantes, notamment dans le cas des AHRS qui font l'objet d'une fusion de donnees.
+Il convient donc d'eliminer ou tout du moins de reduire ces erreurs via une calibration initiale de l'ensemble des capteurs utilises.
+
+\subsection{Types et modele d'erreur}
+
+	Tous les types de capteurs presentent des biais, facteurs d'echelle (scale factors), des erreurs d'inter-couplage (misalignements et cross axis sensitivities) et dans une certaine mesure du bruit aleatoire.
+Des erreurs d'ordre superieur, des erreurs d'inter-couplage d'acceleration angulaire ou encore des erreurs specifiques comme les softs et irons distorsions pour le magnetometre peuvent egalement se produire.
+Pour l'ensemble des capteurs, un modele rendant compte des erreurs lors des differentes mesures peut-etre propose :
+
+\begin{equation}
+	X_i = MS \cdot \left( X_{i,raw} - \textbf{b} \right) + \textbf{n}_i
+\end{equation}
+
+ou $X_i$ est la mesure du capteur comprenant les erreurs, $X_{i, raw}$ la mesure reelle du capteur,  $b$ le biais,  $S$ la matrice rendant compte des scales factor errors, $M$ la matrice des misalignement entre les axes du capteur et $\textbf{n}_i$ le bruit aleatoire avec $i \in \{x,y,z\}$.
+Les differents matrices sont alors \\
+
+\begin{center}
+$S = \begin{pmatrix}  	s_{xx} & s_{xy} & s_{xz} \\ s_{yx} & s_{yy} & s_{yz} \\s_{zx} & s_{zy} & s_{zz}  \end{pmatrix}$ 	
+ et
+ $M = \begin{pmatrix}  1 & -\alpha_{yz} & \alpha_{zy} \\  0 & 1 & -\alpha_{zx} \\  0 & 0 & 1 \end{pmatrix} $
+\end{center}
+
+\vspace{0.2cm}
+
+ou $\alpha_{ij}$ avec $i,j \in \{x,y,z\} $ est l'angle de misalignement entre l'axe $i$ du repere du capteur avec celui de la plateforme selon l'axe $j$ de la plateforme.
+Enfin le biais est represente par un vecteur $\textbf {b} \in \mathbb{R}^3$
+
+\subsection{Methodes adoptees}
+
+Une premiere calibration d'usine est effectuee mais s'avere souvent imprecise ou obsolete du fait de l'usure des capteurs. Des lors, une nouvelle calibration s'avere necessaire.
+
+La calibration des capteurs se fera dans un premier temps independamment les unes des autres et de maniere statique emph{i.
+e.} a l'initialisation.
+Il est a noter que nous ne prendrons pas en compte ici l'influence de la temperature puisqu'il est deja traite dans la calibration d'usine.
+
+\subsubsection{Accelerometre}
+
+Freescale propose une methode de calibration lineaire de l'accelerometre 3 axes.
+Elle repose sur le fait que la somme des amplitudes des composantes normalisees du champ gravitationnel terrestre sur les 3 axes du capteur doit equationetre egale a l'amplitude totale de celui-ci quand le capteur est dans un etat quasi statique.
+Elle consiste a placer l'appareil dans 8 positions donnees pour chaque axe puis d'utiliser une methode d'optimisation de moindres carres pour determiner les 12 parametres de calibration que sont les scale factors et le biais.
+
+Cette methode ne permet toutefois pas de determiner les misalignements.
+
+\subsubsection{Gyroscope}
+
+Zhang  propose une methode de calibration du gyroscope\cite{Gyroscope}.
+Elle repose sur le fait que la somme des mesures selon les 3 axes du gyroscope doit equationetre egale a la vitesse angulaire en entree.
+
+2 etapes sont alors necessaires a la calibration du gyroscope.
+La premiere consiste en des rotations horaires et anti-horaires permettant d'eliminer la vitesse angulaire de rotation de la terre et le biais du gyroscope en prenant le carre de la norme de la difference des integrales des mesures dans les 2 sens puis en utilisant une methode de moindre carre.
+Les scales factors et les angles de misalignements sont ainsi determines.
+Ensuite, les angles de misalignements et les scales factors sont utilises pour estimer le biais via la difference des carres de l'equation de la somme des vitesses angulaires pour 2 positions differentes.
+
+\subsubsection{Magnetometre}
+
+	Vasconcelos et al.
+\cite{MLE} proposent un algorithme dit de maximum likelihood estimation (MLE) qui utilise l'estimation issue de "Two-Step" \cite{Two-Step}, $i.
+e.$ une methode d'estimation de moindre carre suivie d'un filtrage etendu de Kalman (EKF), comme approximation initiale et qui traite le bruit de mesure comme une distribution normale.
+
+	Le probleme de minimisation  la fonction  de vraisemblance (likelihood) qui est quadratique revient alors a determiner les points de l'ellipsoide  qui correspondent au mieux aux mesures du capteur.
+
+	Ce probleme est resolu via la methode du gradient et une methode de Newton descendante pour des espaces euclidiens \cite{Newton} et la regle de Armijo pour la determination de la taille du pas.
+Le gradient et la matrice hessienne de la fonction de log-likelihood sont eux calcules analytiquement.
+
+\subsubsection{Calibration on-board}
+
+	Une calibration on-board peut egalement s'averer necessaire du fait d'un besoin de precision accrue ou de l'obsolescence de la calibration d'usine due a l'usure des capteurs.

### 2016-05-13 Article/state_of_the_art.tex
modif calib and sota tex file

@@
 With $\bar{\epsilon}$ average attitude error (voir \ref{error_definition}) et $\bar{\tau}$ the average computation time. \\
-\begin{figure}[!h]
 
@@
 
-Our framework has been entierely coded in C++ and all tests have been launched on a Dell Optiplex 9020 with Intel $ Core^{TM}$ i7-4790 CPU at 3.60GHz $\times$ 8.
-All curves were generated with \texttt{R} and \texttt{Matlab}.
-No paralelization or code optimisation have been done, our program turns entirely in sequential manner.
-Our goal is to see the difference of computation duration of all methods without modifying the performance with a specific optimisation.
+We  modelled our inertial sensors based on the Fossen's model \cite{fossen_handbook_2011}.
+The noise generated here is Gaussian.
+In this study, sensor bias are not considerate as constant.
+We proceed 30 consecutive simulations in order to get valid results.
+We've used  specification described in table \ref{spec_imu}.
 
-\begin{table}
+\begin{table}[!h]
 \begin{tabular}{|c|c|c|}
@@
 \hline
- Gyroscope & $0.316\times 10^{-2} $  $\mu$rad.s$^{-\frac{1}{2}}$ & $ 3.16$ $ \times 10^{-4} $ $\mu$rad.s$^{-\frac{3}{2}}$ \rule[-1.5pt]{0pt}{13pt}\\
+ Gyroscope & $0.316\times 10^{-4} $  $\mu$rad.s$^{-\frac{3}{2}}$ & $ 0.316$  $\mu$rad.s$^{-\frac{1}{2}}$ \rule[-1.5pt]{0pt}{13pt}\\
  \hline
@@
 
-For our Sensor Model, we've used  specification described in table \ref{spec_imu}.
+All simulations started with an initial attitude error $\delta\mathbi{q}_0$ randomly generated through normal distribution $\mathcal{N}(20,50^2)$.
+The results of score $s_1$ (resp $s_2$) are shown in figure \ref{histo_s1_add} (resp \ref{histo_s2_add}).
+Mean computation times of all methods required for one iteration is given in table (\ref{mean_time}).
 
+\begin{table}[!h]
+\centering
+\begin{tabular}{|c|c|}
+\hline
+Method & Mean Computation time \\
+\hline
+\hline
+CGO & $7.88 \times 10^{-6} s$\rule[-2pt]{0pt}{10pt}\\
+ \hline
+MEKF & $1.12\times 10^{-4} s$\rule[-2pt]{0pt}{10pt}\\
+ \hline
+ USQUE & $1.63\times 10^{-4} s $\rule[-2pt]{0pt}{10pt}\\
+ \hline
+PF & $4.5\times 10^{-2} s$\rule[-2pt]{0pt}{10pt}\\
+ \hline
+REQUEST& $1.91\times 10^{-4} s$ \rule[-2pt]{0pt}{10pt} \\
+ \hline
+\end{tabular}
+\caption{Mean computation times for one iteration}
+\label{mean_time}
+\end{table}
+
+\begin{figure}[!h]
 \begin{subfigure}{.5\textwidth}

### 2016-05-16 Article/state_of_the_art.tex
update sota

@@
 For attitude representation, the quaternion must satisfies a single constraint given by  $|| \mathbi{q}  || = 1$.
-
-Then, the attitude matrix $R \in SO(3) $ related to the quaternion $\mathbi{q}$ is given by :
+The attitude matrix $R \in SO(3) $ related to the quaternion $\mathbi{q}$ is given by :
 \vspace{-0.15cm}
@@
 Where $SO(3)$ is the special orthogonal group defined by:
+
 \vspace{-0.15cm}
@@
 The Kalman filter provides an optimal estimation for linear systems.
-Nevertheless, most of studied dynamical systems are nonlinear (attitude estimation, pendula, ...
-).
+Nevertheless, most of studied dynamical systems are nonlinear (like attitude estimation problem).
 Thus, extended Kalman Filter approach  was proposed for nonlinear systems \cite{larson1967application, larson1967precomputation,terejanu2008extended}.
@@
 \cite{mahony_nonlinear_2008} have proposed an nonlinear observer called Constant Gain Observer (CGO), termed the explicit complementary filter, that provides attitude estimates as well as gyro bias estimates.
-
 One problem of Kalman Filters is to find the initial values in the covariance (matrix $Q$) because it is difficult to relate it to real physical quantities.
 The covariance matrix is not involve in nonlinear observer theory.
-This is one of the reasons there has been an increasing interest for nonlinear observer design the last decades.
+This is one of the reasons there has been an increasing interest for nonlinear observer design during the last decades.
 Another reason is their low computational need.
@@
 
-The attitude determination problem for vector observation was first formulated as a least squares estimation problem by Wahba \cite{wahba_least_1965} in 1965 :\\
+The attitude determination problem for vector observation was first formulated as a least squares estimation problem by Wahba \cite{wahba_least_1965} in 1965 :
+\vspace{0.1cm}
 
@@
 
-The matrix $R(\mathbi{q})$, is the rotation matrix that corresponds to the quaternion $\mathbi{q}$ (equation \ref{quat_to_rot}). The QUEST algorithm are conducted in the following manner. Let
+The matrix $R(\mathbi{q})$, is the rotation matrix that corresponds to the quaternion $\mathbi{q}$ (equation \ref{quat_to_rot}).
+The Quatenion Estimator (QUEST) algorithm propose a solution to the Wahba's problem.
+Let :
 
@@
 
-The minimisation problem from Equation (\ref{QUEST_bis}) is now a maximisation problem.
-thus, the solution $\mathbi{q}_{opt}$ is the eingen vector correspond to the largest eigen value of $K$ \cite{markley1999estimate}.
+The minimization problem from equation (\ref{QUEST_bis}) is now a maximization problem.
+thus, the solution $\mathbi{q}_{opt}$ is the eingen vector correspond to the highest eigen value of $K$ \cite{markley1999estimate}.
 The main problem of QUEST is that it does not take into account of all the previous observations performed since the beginning.
-A recursive QUEST algorithm (REQUEST) proposed by Bar-Itzhack manages this problem.
-Contrary to other methods, the QUEST algorithms do not estimate the gyroscope bias.
+Then, a REcursive-QUEST algorithm (REQUEST) proposed by Bar-Itzhack\cite{bar-itzhack_request-recursive_1996} manages this problem.
+Contrary to other methods, solutions to Wahba's problem do not estimate the gyroscope bias.
 Which can make it less robust in some cases.
-
 There are also other methods based on Wahba's problem like the Davenport's q-method \cite{weighted1971nasa} proposed in 1971 considered more robust than QUEST.
 Or the method ESOQ proposed by Mortari \cite{mortari1997esoq}  in 1997.
-However, for this library, we limit our investigation to REQUEST.\\
+However, for this library, we limit our investigation to REQUEST.
 
@@
 
-Particle filtering method is based on Monte-Carlo method\cite{metropolis1949monte}.
-The main idea is to use a large number of samples called particles to estimate a nonlinear function \cite{chen_bayesian_2003} and \cite{terejanu2009tutorial}.
+Particle filtering method is based on Monte-Carlo simulation\cite{metropolis1949monte}.
+The main idea is to use a large number of samples called particles to estimate a nonlinear function \cite{chen_bayesian_2003,terejanu2009tutorial}.
 The set of $N$ particle with associated weight at time $k$ are denoted by $\{\textbf{x}^{(i)}_{k},w^{(i)}_{k}\}$.
@@
 We have also considered the revisited algorithm of Cheng \cite{chang_particle_2014} based on bootstrap filter\cite{gordon1993novel}.
-We have chosen to implement this one because they use the same approach as USQUE \cite{crassidis_unscented_2003} and MEKF\cite{markley2003attitude}.
-We decided to explore the performance of this three algorithms with the same model.
-The particle filter provides an estimate of the nonlinear system without assuming that the noise is Gaussian.
-So it is more robust to any situation, which makes it more performant in the context of a real applications.
-But an important computation time is required and that is a limitation for a real-time application.
+We have chosen to implement this one because they use the same approach as USQUE  and MEKF.
+The particle filter does not assume that the measurement noise is Gaussian.
+Which makes it more adaptable in the context of a real applications.
+Nevertheless an important computation time is required.
 
@@
 \subsection{Score definition}
-We have access through our designed ground truth (simulated or from computer vision technics) to the true attitude $\mathbi{q}^{real}_k$ at time $k$ of the reference frame $\{c\}$ related to $\{a\}$.
-By computation, each method returns an estimate attitude $\mathbi{q}^{est}_k$ at time $k$ (see figure \ref{test_method}).
+We have access through the ground truth (simulated or from computer vision technics) to the true attitude $\mathbi{q}^{real}_k$ at time $k$ of the reference frame $\{c\}$ related to $\{a\}$.
+By computation, each method returns an estimate attitude $\mathbi{q}^{est}_k$ at time $k$.
 
@@
 
-With $n$ the number of data at our disposal. We use this information (mean error) for our performance study.
+With $n$ the number of iteration at our disposal.
+We use this information (mean error) for our performance study.
+To differentiate the methods, we assigned  some scores in order to highlight the behaviour of these methods in different configurations defined as:
 
-\begin{figure}[!h]
-\centering
-\includegraphics[scale=0.40]{images/test_method.png}
-\caption{Algorithm designed to compute attitude estimation methods' performances}
-\label{test_method}
-\end{figure}
-
-To differentiate the methods, we assigned  some ''scores'' in order to highlight the behaviour of these methods in different configurations. We defined these ''scores'' as:
+%\begin{figure}[!h]
+%\centering
+%\includegraphics[scale=0.40]{images/test_method.png}
+%\caption{Algorithm designed to compute attitude estimation methods' performances}
+%\label{test_method}
+%\end{figure}
 
@@
 
-With $\bar{\epsilon}$ average attitude error (voir \ref{error_definition}) et $\bar{\tau}$ the average computation time. \\
+With $\bar{\epsilon}$ average attitude error (see \ref{error_definition}) et $\bar{\tau}$ the average computation time. \\
 
@@
 In this study, sensor bias are not considerate as constant.
-We proceed 30 consecutive simulations in order to get valid results.
-We've used  specification described in table \ref{spec_imu}.
+We proceed 30 consecutive simulations in order to get valid results with data performance specification described in table \ref{spec_imu}.
 
@@
 
-All simulations started with an initial attitude error $\delta\mathbi{q}_0$ randomly generated through normal distribution $\mathcal{N}(20,50^2)$.
+All simulations started with an initial attitude error $\delta\mathbi{q}_0$ randomly generated through normal distribution.
 The results of score $s_1$ (resp $s_2$) are shown in figure \ref{histo_s1_add} (resp \ref{histo_s2_add}).
@@
  \hline
+REQUEST& $1.91\times 10^{-4} s$ \rule[-2pt]{0pt}{10pt} \\
+ \hline
 PF & $4.5\times 10^{-2} s$\rule[-2pt]{0pt}{10pt}\\
  \hline
-REQUEST& $1.91\times 10^{-4} s$ \rule[-2pt]{0pt}{10pt} \\
- \hline
 \end{tabular}
-\caption{Mean computation times for one iteration}
+\caption{Mean computation times for one iteration according to the methods available in the library}
 \label{mean_time}
 \end{table}
+
+The library has been entierely coded in C and all tests have been launched on a Dell Optiplex 9020 with Intel $ Core^{TM}$ i7-4790 CPU at 3.60GHz $\times$ 8.
+All curves were generated with \texttt{R} and \texttt{Matlab}.
+No paralelization or code optimization have been done, our program turns entirely in sequential manner.
+Some optimization will come in the next versions.
 
@@
 We have designed a system that can be easily repeatable in laboratory.
-We used a smartphone as our IMU and a calibrated webcam (see figure \ref{situation_validation} ).
-With this system, we have defined our ground-truth as follows.
-A marker is stuck on the smartphone and we calculate an estimate of the attitude of the marker (which represents the smartphone) relative to the reference frame of the webcam $\{w\}$.
-We used the library \texttt{ArUco}  as optical feedback.
+We used a smartphone as our IMU and a calibrated webcam (see figure \ref{situation_validation}).
+With this system, we have defined our ground-truth as follows: Some markers are stuck on the smartphone and we calculate an estimate of the quaternion attitude $\mathbi{q}_c^w$ of the markers (which represents the smartphone) relative to the reference frame of the webcam $\{w\}$.
+To do so, we've used the library \texttt{ArUco} \cite{garrido2014automatic} as optical feedback.
 At the same time, we get the data from the IMU of the smartphone to proceed the attitude estimation with our different algorithms.
+At time $t = 0$, the camera and the smartphone have been set such as attitude quaternions $\mathbi{q}^a_w$, $\mathbi{q}^c_a$ et $\mathbi{q}^w_c$ are approximatively known.
+The camera being fixed, $\mathbi{q}^a_w $ does not vary.
+Moreover, from figure \ref{situation_validation}  we have:
 
@@
 \end{figure}
-
-At time $t = 0$, the camera and the smartphone has been set such as attitude quaternions $\mathbi{q}^a_w$, $\mathbi{q}^c_a$ et $\mathbi{q}^w_c$ are approximatively known.
-The camera being fixed, $\mathbi{q}^a_w $ does not vary.
-Moreover, from figure \ref{situation_validation}  we have:
 
@@
 
-All methods give an estimate of $\mathbi{q}^c_a$ , that are compared with the data from the camera.
-The optical feedback provides a ground truth estimate for this experiment.
+All methods give an approximation of $\mathbi{q}^c_a$ , that are compared with the data from the camera.
 To estimate the error that can occur with the optical feedback.
@@
 Then, another rotation of 90 degree around the $z_c$ on the smartphone but in anticlockwise direction to return to the initial state.
-In order to study possible divergence of all methods.
+In order to study possible divergence.
+The results is shown in figure ...

### 2016-05-20 Article/state_of_the_art.tex
modif results simulation

@@
 
-The library has been entierely coded in C and all tests have been launched on a Dell Optiplex 9020 with Intel $ Core^{TM}$ i7-4790 CPU at 3.60GHz $\times$ 8.
-All curves were generated with \texttt{R} and \texttt{Matlab}.
-No paralelization or code optimization have been done, our program turns entirely in sequential manner.
-Some optimization will come in the next versions.
-
 \begin{figure}[!h]
@@
 \caption{Scores $s_1$ (accuracy) and $s_2$ (accuracy and computation time ratio) of methods where noise is AWGN with 3 strength (weak, moderate, strong)}
+\end{figure}
+
+From figures \ref{histo_s1_add} et \ref{histo_s2_add}, we can observe that USQUE has the best score $s_1$, which proove its effectiveness in term of accuracy.
+While CGO has the highest $s_2$, that make it more interesting for embedded system context.
+On a other hand, it seems that PF is not as accurate as expected.
+This is due to the fact that its covariance time requirement is more important than USQUE.
+Moreover, as shown in figure \ref{error_plot}, even when the noise is strong, PF become more reliable than USQUE after a fex second of utilization.
+\\
+
+The library has been entierely coded in C and all tests have been launched on a Dell Optiplex 9020 with Intel $ Core^{TM}$ i7-4790 CPU at 3.60GHz $\times$ 8.
+All curves were generated with \texttt{R} and \texttt{Matlab}.
+No paralelization or code optimization have been done, our program turns entirely in sequential manner.
+Some optimization will come in the next versions.
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.3]{images/error_plot.png}
+\caption{Plot of the RMS attitude error (score $s_1$)over time when the noise is AWGN with strong strength.
+Here only PF, USQUE and CGO is represented.
+We can observe that PF and CGO require more time to converge to the solution.
+However, after a few second, PF become more reliable than USQUE.
+}
+\label{error_plot}
 \end{figure}

### 2016-08-04 Article/conclusion.tex
Modification Article, Calibration file + Introduction + Conclusion

@@
 
-notre library est trop bien...
+In this paper, we presented our open source library \texttt{OpenMotion} and its different functionalities.
+Our aim is to make it a powerful tool for developing application or new algorithms.
+Combining this library with our comparative framework \cite{braudcomparison} could be an interesting tool for research.
+A performance study has been processed too in order to show the capacities of the algorithms present in the library.
+Thus, the user has access to the algorithm that fit the best to his needs and the open source feature allows him to have a better understanding of the algorithms.
+For future work, we continue to develop the library and focus on his robustness and adaptability to every type of smartphone.
+With the developing sensor calibration, better accuracy is expected too.
 

### 2016-08-04 Article/introduction.tex
Modification Article, Calibration file + Introduction + Conclusion

@@
 \section{Introduction}
+
+%context + motivation
+
+Since the last decades, it is easily observable that smartphones market has grown drastically.
+Indeed, this devices propose convenient functionalities for household.
+On the other hands, the different kind of sensor available and easily accessible make smartphones an interesting tool for research.
+Nowadays, some applications where smartphone can be used attract the attention of researchers such as Indoor Navigation\cite{huang2009survey}.
+Severals solution for this has been proposed with different approaches like using only computer vision technics \cite{desouza2002vision,bonin2008visual} or wireless networks\cite{gu2009survey}.
+Another approaches is to sense the motion by using some inertial sensors.
+Among its sensor and according to the smartphone, an Inertial Measure Unit (IMU) is vacant and used when coupling with a magnetometer as an Attitude and Heading Reference System (AHRS).
+However, an AHRS sensor may be not robust or accurate in certain situation due to unpredictable event or sensor deficiencies.
+Moreover, for Indoor Localization or Navigation applications, it is crucial to have an algorithm that propose an optimal estimation of the attitude in term of accuracy and computation time requirement.
+This is the reason why researchers are interested in sensor fusion approaches because it overcomes some problems such as sensor drifting.
+
+Since 1960, severals data fusion has been developed after a seminal framework on data fusion has been initiated by Kalman and  Bucy \cite{kalman_new_1960}.
+In this paper, the most common family of data fusion algorithm is presented briefly.
+However, implementing this kind of algorithm require time and resources.
+Furthermore, the available software of data fusion remains general and may be not adaptable for the inertial case.
+For this reason, we propose an open source and free library called \texttt{OpenMotion} that contains different family of sensor fusion algorithms designed to fit to the inertial case.
+For this paper, the different functionalities proposed by the library are presented.
+In a simulated environment, a performance study of all methods has been processed and the values are shown.
+To quantify the performances, we determined 2 kind of scores,  $s_1$ representing the accuracy of the method and $s_2$ that takes into account the accuracy and the computation time.
+The $s_2$ give better visibility of the methods' effectiveness for usage in embedded devices, which is interesting for smartphone application.
+Furthermore, working only in a simulated environment is not enough to have a good visibility.
+Moreover, in our context, it is very difficult to obtain a reliable ground truth.
+For this study, we designed the ground truth  based on computer vision methods.

### 2016-08-04 Article/prez_library.tex
Modification Article, Calibration file + Introduction + Conclusion

@@
 Our approach is based on data fusion technics using only the output of a IMU (composed by a 3D gyroscope,  3D accelerometer and a 3D magnetometer).
-Among several methods proposed by the library, the user has the possibilities to choose which methods fits the best to his expectation according to its performance.
-Here the 2 reference frame to remember.
+Among several sensor fusion algorithms proposed by the library, the user has the possibilities to choose the one that fits the best to his expectation according to its performance.
+
+However, we designed the library also as an academic tool.
+For this reason, we have chosen to implement for this library the most common family of methods.
+In previous works \cite{braudcomparison}, we proposed a comparative framework that allow us to observe the performance of this different methods in a quantitative manner.
+Thus, the user has in his disposition the code and and the performance study.
+This will help him to have a better understanding of the algorithms' mechanism and behavior.
+On the other hand, the user can use the library as a developing tool to design and create new sensor fusion algorithm.
+Moreover, he can compare his work directly with the available methods.
+For this paper,  2 references frames are used as follow:
 
@@
 \texttt{OpenMotion} provide an estimation of the orientation of $\{c\}$ relative to $\{a\}$.
-The choice of the output type (quaternion, rotation matrix, euler angle) is according to the user needs.
+The choice of the output type (quaternion, rotation matrix, euler angle) is also according to the user needs.
 
@@
 \includegraphics[scale=0.65]{images/Schema_situation.png}
-\caption{Definition of the scene with 2 frame:  a fixed frame NED (North East Down) noted $\{a\}$ and a moving frame object noted $\{c\}$}
+\caption{Definition of the scene with 2 frame:  a fixed frame NED (North East Down) noted $\{a\}$ and a moving frame object noted $\{c\}$.
+\texttt{OpenMotion} will provide an estimate of the attitude of $\{c\}$ relative to $\{a\}$.
+}
 \label{Schema_situation}
 \end{figure}
-
-%
-%\begin{lstlisting}[caption=Exemple of utilisation in C++,label=code]
-%// om_exemple.cpp
-%#include <om.h>
-%
-%using namespace om;
-%
-%void main(int argc, char** argv){
-%
-%	// 
-%	SensorFusionManager manager;
-%
-%	//if calibration is necessary
-%	manager.calibration();
-%	
-%	// while IMU send data
-%	while( isData ){
-%
-%		// set all IMU data at time t
-%		manager.setGyroscopeData(gyro_x,gyro_y,gyro_z);	
-%		manager.setAccelerometerData(acc_x,acc_y,acc_z);
-%		manager.setMagnetometerData(mag_x,mag_y,mag_z);
-%		
-%		// get different kind of output according to user's choices
-%		
-%		// quaternion
-%		Quaternion q_est = manager.getAttitude<Quaternion>(); 
-%		
-%		//  rotation matrix
-%		Matrix R = manager.getAttitude<Matrix>(); 
-%		
-%		//  axis angle
-%		AxisAngle axisAngle = manager.getAttitude<Matrix>(); 
-%		...
-%	}
-%}
-%
-%\end{lstlisting}
 

### 2016-08-04 Article/state_of_the_art.tex
Modification Article, Calibration file + Introduction + Conclusion

@@
 The Quatenion Estimator (QUEST) algorithm propose a solution to the Wahba's problem.
-Let :
-
-\begin{equation}
-B = \sum_{i=1}^N a_i\textbf{b}_i \textbf{r}_i^T
-\end{equation}
-
-Then, we can rewrite the equation (\ref{QUEST}) as follow
-
-\begin{equation}
-J(\mathbi{q}) = \frac{1}{2} \sum_{i=1}^n a_i  - tr(R(\mathbi{q})B^T)
-\label{QUEST_bis}
-\end{equation}
-
-Due to the fact that the representation of the attitude matrix is a homogenous quadratic function of $\mathbi{q}$, we can say that
-
-\begin{equation}
-tr(R(\mathbi{q})B^T) = \mathbi{q}^TK\mathbi{q}
-\label{QUEST_ter}
-\end{equation}
-
-where $K$ is the symmetric traceless matrix such as
-
-\begin{equation}
-K = \begin{pmatrix} S-tr(B)I_3 & \textbf{z} \\ \textbf{z}^T & tr(B)
-\end{pmatrix}
-\end{equation}
-
-with
-
-\begin{equation}
-\left\{\begin{array}{l}
-S = B + B^T\\
-\textbf{z} = \sum_ia_i\textbf{b}_i\times\textbf{r}_i
- \end{array}
-\right.
-\end{equation}
-
-The minimization problem from equation (\ref{QUEST_bis}) is now a maximization problem.
-thus, the solution $\mathbi{q}_{opt}$ is the eingen vector correspond to the highest eigen value of $K$ \cite{markley1999estimate}.
 The main problem of QUEST is that it does not take into account of all the previous observations performed since the beginning.
@@
 We use this information (mean error) for our performance study.
-To differentiate the methods, we assigned  some scores in order to highlight the behaviour of these methods in different configurations defined as:
-
-%\begin{figure}[!h]
-%\centering
-%\includegraphics[scale=0.40]{images/test_method.png}
-%\caption{Algorithm designed to compute attitude estimation methods' performances}
-%\label{test_method}
-%\end{figure}
+To differentiate the methods, we assigned  some scores in order to highlight the behavior of these methods in different configurations defined as:
 
@@
 \end{tabular}
-\caption{Calibrated IMU data performance specification, Gyrometer, Accelerometer, Magnetometer}
+\caption{Calibrated IMU data performance specification, Gyroscope, Accelerometer, Magnetometer}
 \label{spec_imu}
@@
 The results is shown in figure ...
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.35]{images/situation_validation.png}
+\caption{Representation of different reference frame used in this experiment. Here, only $ \text{\mathbi{q}}_w^a $  does not change.}
+\label{situation_validation}
+\end{figure}

### 2016-08-05 Article/Smartphone/calibration.tex
Modify Article folder and Readme for MAC OS installer

\section{Calibration and initialization}

%	Les systemes embarque comme les smartphones disposent pour la plupart de capteurs servant au systeme ou a l'utilisateur pour se renseigner sur des grandeurs utiles quant a l'appareil lui-meme ou bien a son environnement. Ces capteurs sont donc varies mais on peut notamment distinguer les capteurs constituant une centrale inertielle ou IMU pour Inertial Measurement Unit que sont les accelerometres et les gyroscopes 3 axes. Ces capteurs sont de plus en plus associes avec un magnetometre. Ainsi, ces capteurs peuvent constituer un AHRS pour Attitude and Heading Reference System ayant pour but de reperer l'appareil dans l'espace 3D. L'accelerometre sert ici a mesurer l'acceleration lineaire, le gyroscope la vitesse angulaire et le magnetometre le champ magnetique terrestre local et sert dans la plupart des cas de compas. Dans le cas d'un smartphone, ces capteurs sont low-cost et de type MEMS pour Micro-Electro-Mechanical Systems. Des lors, le signal utile etant faible, l'erreur induite peut-etre du meme ordre de grandeur et donc conduire a des imprecisions importantes, notamment dans le cas des AHRS qui font l'objet d'une fusion de donnees. Il convient donc d'eliminer ou tout du moins de reduire ces erreurs via une calibration initiale de l'ensemble des capteurs utilises.

Embedded system such as smartphones has in disposition several kind of sensors that give some values about the environment or the system to the user. Among this sensor and according to the smartphone, a IMU is provided. However, in the case of smartphones, these sensors are low-cost. Thus, the useful signal being low, the error induced may be of the same order of magnitude and therefore lead to significant inaccuracies. It is therefore necessary to minimize these errors using an initial calibration of all sensors used.

\subsection{Error Model}

%	Tous les types de capteurs presentent des biais, facteurs d'echelle (scale factors), des erreurs d'inter-couplage (misalignements et cross axis sensitivities) et dans une certaine mesure du bruit aleatoire. Des erreurs d'ordre superieur, des erreurs d'inter-couplage d'acceleration angulaire ou encore des erreurs specifiques comme les softs et irons distorsions pour le magnetometre peuvent egalement se produire.
%Pour l'ensemble des capteurs, un modele rendant compte des erreurs lors des differentes mesures peut-etre propose :

Every sensors has a bias, scale factors, misalignments, cross axis sensitivities  and random measurement disturbances.  Errors of higher order, the inter-coupling error, angular acceleration or the specific errors such as soft and iron distortions from the magnetometer may also occur.

For all sensors, a model  the errors of the different measures can be proposed as follow:

\begin{equation}
	X_i = MS \cdot \left( X_{i,raw} - \textbf{b} \right) + \textbf{n}_i
\end{equation}

%
%ou $X_i$ est la mesure du capteur comprenant les erreurs, $X_{i, raw}$ la mesure reelle du capteur,  $b$ le biais,  $S$ la matrice rendant compte des scales factor errors, $M$ la matrice des misalignement entre les axes du capteur et $\textbf{n}_i$ le bruit aleatoire avec $i \in \{x,y,z\}$.
%Les differents matrices sont alors \\

Where  $X_i$ is the sensor's output that include errors, $X_{i, raw}$ is the true signal, $\textbf{b}$ is the sensor's bias, $S$ is the scale factor matrix, $M$ is the misalignments matrix and $\textbf{n}_i$  some noise. Matrices $S$ and $M$ are defined as follow \\

\begin{center}
$S = \begin{pmatrix}  	s_{xx} & s_{xy} & s_{xz} \\ s_{yx} & s_{yy} & s_{yz} \\s_{zx} & s_{zy} & s_{zz}  \end{pmatrix}$ 	
 et
 $M = \begin{pmatrix}  1 & -\alpha_{yz} & \alpha_{zy} \\  0 & 1 & -\alpha_{zx} \\  0 & 0 & 1 \end{pmatrix} $
\end{center}

\vspace{0.25cm}

A REVOIR ICI\\
est l'angle de misalignement entre l'axe $i$ du repere du capteur avec celui de la plateforme selon l'axe $j$ de la plateforme. Enfin le biais est represente par un vecteur $\textbf {b} \in \mathbb{R}^3$

\subsection{Adopted methods}

%Une premiere calibration d'usine est effectuee mais s'avere souvent imprecise ou obsolete du fait de l'usure des capteurs. Des lors, une nouvelle calibration s'avere necessaire.
%
%La calibration des capteurs se fera dans un premier temps independamment les unes des autres et de maniere statique emph{i. e.} a l'initialisation. Il est a noter que nous ne prendrons pas en compte ici l'influence de la temperature puisqu'il est deja traite dans la calibration d'usine.

A first calibration made in factory is performed but it is inaccurate or obsolete due to sensor wear. Thus, an additive calibration is required. Sensor calibration will at first be processed independently from each other in a static manner. For this paper, the calibration do not take into account here the influence of temperature because it is already treated in the factory calibration.

\subsubsection{Accelerometer}

%Freescale propose une methode de calibration lineaire de l'accelerometre 3 axes. Elle repose sur le fait que la somme des amplitudes des composantes normalisees du champ gravitationnel terrestre sur les 3 axes du capteur doit equationetre egale a l'amplitude totale de celui-ci quand le capteur est dans un etat quasi statique. Elle consiste a placer l'appareil dans 8 positions donnees pour chaque axe puis d'utiliser une methode d'optimisation de moindres carres pour determiner les 12 parametres de calibration que sont les scale factors et le biais.
%
%Cette methode ne permet toutefois pas de determiner les misalignements.

Freescale(citation ici) propose a linear calibration method for 3-axis accelerometer. It is based on the fact that the sum of amplitude of the earth's gravitational field normalized component  on the 3-axis sensor must be equal to the total of amplitudes when the sensor is in a quasi-static state. the device is placed  in 8 data positions for each axis and then a optimization method of least squares is performed to determine the 12 calibration parameters that are factoring scale and bias. However, this method does not allow to determine the misalignments.

\subsubsection{Gyroscope}

%Zhang  propose une methode de calibration du gyroscope\cite{Gyroscope}. Elle repose sur le fait que la somme des mesures selon les 3 axes du gyroscope doit equationetre egale a la vitesse angulaire en entree. 
%
%2 etapes sont alors necessaires a la calibration du gyroscope. La premiere consiste en des rotations horaires et anti-horaires permettant d'eliminer la vitesse angulaire de rotation de la terre et le biais du gyroscope en prenant le carre de la norme de la difference des integrales des mesures dans les 2 sens puis en utilisant une methode de moindre carre. Les scales factors et les angles de misalignements sont ainsi determines. Ensuite, les angles de misalignements et les scales factors sont utilises pour estimer le biais via la difference des carres de l'equation de la somme des vitesses angulaires pour 2 positions differentes.

Zhang  propose a calibration method for the gyroscope based on the idea that the sum of the measures must be equals to the input angular velocity.  Then, 2 step is necessary. The first executes clockwise rotation and anti-clockwise rotation such as angular velocity from earth rotation and the bias are minimized.

2 etapes sont alors necessaires a la calibration du gyroscope. La premiere consiste en des rotations horaires et anti-horaires permettant d'eliminer la vitesse angulaire de rotation de la terre et le biais du gyroscope en prenant le carre de la norme de la difference des integrales des mesures dans les 2 sens puis en utilisant une methode de moindre carre. Les scales factors et les angles de misalignements sont ainsi determines. Ensuite, les angles de misalignements et les scales factors sont utilises pour estimer le biais via la difference des carres de l'equation de la somme des vitesses angulaires pour 2 positions differentes.

\subsubsection{Magnetometer}

%	Vasconcelos et al. \cite{MLE} proposent un algorithme dit de maximum likelihood estimation (MLE) qui utilise l'estimation issue de "Two-Step" \cite{Two-Step}, $i. e.$ une methode d'estimation de moindre carre suivie d'un filtrage etendu de Kalman (EKF), comme approximation initiale et qui traite le bruit de mesure comme une distribution normale.
%
%	Le probleme de minimisation  la fonction  de vraisemblance (likelihood) qui est quadratique revient alors a determiner les points de l'ellipsoide  qui correspondent au mieux aux mesures du capteur.
%
%	Ce probleme est resolu via la methode du gradient et une methode de Newton descendante pour des espaces euclidiens \cite{Newton} et la regle de Armijo pour la determination de la taille du pas. Le gradient et la matrice hessienne de la fonction de log-likelihood sont eux calcules analytiquement.

Vasconcelos et al. \cite{MLE}  propose a maximum likehood estimation (MLE) method. They use the output of Two-Step \cite{Two-Step} as initial approximation. Furthermore , they assume a normal distribution for the measurement noise.

	Le probleme de minimisation  la fonction  de vraisemblance (likelihood) qui est quadratique revient alors a determiner les points de l'ellipsoide  qui correspondent au mieux aux mesures du capteur.

	Ce probleme est resolu via la methode du gradient et une methode de Newton descendante pour des espaces euclidiens \cite{Newton} et la regle de Armijo pour la determination de la taille du pas. Le gradient et la matrice hessienne de la fonction de log-likelihood sont eux calcules analytiquement.

\subsubsection{On-board calibration}

%	Une calibration on-board peut egalement s'averer necessaire du fait d'un besoin de precision accrue ou de l'obsolescence de la calibration d'usine due a l'usure des capteurs.

On-board calibration can also be necessary when a high accuracy is needed or the factory calibration is obsolete.
